### 2021-03-15 papers/SYRCoSE/intro.tex
SYRCoSE: move intro into separate folder

\section{Introduction}

\subsection{Motivation}

% Терия структур событий (СС) была придумана Винскелем в 19?? году [статья]. С тех пор она активно развивалась, в различных статьях рассматривалось как различные подходы к этой теории: в [этих статьях] рассмотриваласть категорная семантика, в [этих статьях] [какая-то другая], так и применение СС можно в различных проложениях.

% Для каких же приложений можно использовать СС теорию? Например, для задания формальной семантики конкуретных систем [ссылки]. Истина познается в сравнении, так что давайте попробуем понять неформальный смысл СС на примере задания семнатики какой-нибудь параллельной программы

% Рассмотрим следущую параллельную программу
\begin{tikzpicture}

      \node (iwy) %[left of = mid, node distance = 1.1cm] 
      at (-1.3, 2.5)
      {$x:=1;$};

      \node (irx) [above = -0.5cm of iwy.west, anchor = west] 
      {$r:=x;$};

      \node (iwx) %[right of = mid, node distance = 0.8cm] 
       at (0.5, 2.5)
      {$x:=2;$};

      \draw ($ (iwy.south east) + (0.1,-0.5) $) -- ($ (iwy.south east) + (0.1,0.6) $);
      \draw ($ (iwx.south west) + (-0.15,-0.5) $) -- ($ (iwx.south west) - (0.15,-0.6) $);
\end{tikzpicture}
% Программа разделяется на два потока: в первом мы пишем в х 1, потом читаем из него что-то в r, а во втором пишем в х 2. Вопрос: что окажется в переменной r после исполнения программы?
% Если мы, например, выполним сначала инструкции из первого потока, а потом из второго то в r окажется 1. Такое поведение отражается на следующей диаграмме:

\begin{tikzpicture}
    \node (ts1) at (-3,2) {ThreadStart};
    \node (ts2) at (-1, 2) {ThreadStart};
    \node (wxo) at (-3,0.2) {$\wlab{x}{1}$};
    \node (wxt) at (-1,0.2) {$\wlab{x}{2}$};
    \node (rxo) at (-3, -2.2) {$\rlab{x}{1}$};
    \draw[ca] (wxo)  -- node{$\leqslant\qquad$}  (ts1);
    \draw[ca] (wxt)  -- node{$\leqslant\qquad$}  (ts2);
    \draw[ca] (rxo)  -- node{$\leqslant\qquad$}  (wxo);
\end{tikzpicture}

% В этой диаграмме мы следим за тем, что происходит с переменной x. Тут есть несколько типов событий:
% - W(x,u) означает что мы записали в x u
% - R(x,u) означает что мы прочитали из x u
% - ThreadStart означает начало потока
% Так же на этой диаграмме есть отношение $\leq$ -- причинно следственная связь между собыитиями: например оно связывает W(x,1) и R(x,1), что логично, так как запись в x 1 повлекла чтение из x 1 (формально в 2A)
% Но возможно и другое исполнение программы в котором из x мы прочитаем 2. Ему соотвесвует другая диаграмма:
\begin{tikzpicture}
    \node (ts1) at (-3,2) {ThreadStart};
    \node (ts2) at (-1, 2) {ThreadStart};
    \node (wxo) at (-3,0.2) {$\wlab{x}{1}$};
    \node (wxt) at (-1,0.2) {$\wlab{x}{2}$};
    \node (rxt) at (-3, -2.2) {$\rlab{x}{2}$};
    \draw[ca] (wxo)  -- node{$\leqslant\qquad$}  (ts1);
    \draw[ca] (wxt)  -- node{$\leqslant\qquad$}  (ts2);
    \draw[ca] (rxt)  -- node{$\leqslant\qquad$}  (wxo);
    \draw[ca] (rxt)  -- node{$\leqslant\qquad$}  (wxt);
\end{tikzpicture}
% Заметим что, то тут образовалась новая стрелка из W(x,2) в R(x,2), так как чтение из x 2 не могло бы произойти еслиб мы туда 2 не записывали. Однако стрелка из W(x,1) тоже не исчезла, потому что, $x := 1$ в программе стоит перед $r := x$, а значит мы не можем выполнить $r := x$ не выполнив перед этим $x := 1$.
%  Тут мы видим, что при чтении их x мы столкнулись с некоторой недетерменнированностью, которая отразилось в том, что у нас появилось целое множество графоф, соотвествующих разным исполнения программы
% По множеству всех таких возможных графов можно понять все возможные поведения программы, однако с ростом недетеменированности в каком-то одном месте растет количество графов, которые отличаются друг от друга только там. Хочется использовать такой формализм, который позволил бы описывать локальный недетерменизм более компактно.   
% Здесь нам как раз может помочь второе отношение, присутсвующее в структурах собыий — это отношение конфликта между событиями: e1 # e2. 
% Оно обозначает, что e1 и e2 не могут произойти в одном исполнении программы например R(x,2) и R(x,1) из примера выше. За счет этого в одной СС можно закодировать сразу несколько иполнений программы [статьи], а так же моделировать исполнения, в которых, в ходе оптимизаций, произошли перестановки различных интрукций [про LB и es]. 
% Вот так кодируется наш пример с использованием отношкния конфликта
\begin{tikzpicture}
    \node (ts1) at (-3,2) {ThreadStart};
    \node (ts2) at (-1, 2) {ThreadStart};
    \node (wxo) at (-3,0.2) {$\wlab{x}{1}$};
    \node (wxt) at (-1,0.2) {$\wlab{x}{2}$};
    \node (rxt) at (-2, -2.2) {$\rlab{x}{2}$};
    \node (rxo) at (-4, -2.2) {$\rlab{x}{1}$};
    \draw[ca] (wxo)  -- node{$\leqslant\qquad$}  (ts1);
    \draw[ca] (wxt)  -- node{$\leqslant\qquad$}  (ts2);
    \draw[ca] (rxt)  -- node{$\leqslant\qquad$}  (wxo);
    \draw[ca] (rxt)  -- node{$\leqslant\qquad$}  (wxt);
    \draw[ca] (rxo)  -- node{$\leqslant\qquad$}  (wxo);
    \draw[cf] (rxo)  -- node[above]{$\#$}  (rxt);
\end{tikzpicture}

\subsection{Problem Statement}

% В существующих работах по структурам событий обычно результаты либо не формализованы ни в каких пруф-ассистантах [ссылки], либо слишком привязаны к конкретной предметной области или задаче [ссылки], что затрудняет переиспользование теории структур событий в других проектах [расписать про это в Related Works].

%Нашей задачей стало решить две, описанные выше, проблемы \\ 1) Формализовать теорию СС в каком-нибудь proof-assistance: предоставить леммы и тактики упрощаюшие работу с СС и доказать основные результаты, описанные в статьях

% \\ 2) Показать, как нашу библиотеку можно использовать для решения прикладных задач, например для задания  семантики абстрактной регистровой машины с слабо-согласованной моделью разделяемой памяти.

%\subsection{Outline} %описание статьи

### 2021-03-16 papers/SYRCoSE/intro.tex
SYRCoSE: reworking the introduction

@@
 \section{Introduction}
+
+Event structures is a mathematical formalism introduced 
+by Winskel~\cite{Winskel:86} as a semantic domain of concurrent programs.
+In recent years there is a renewed interest in event structures, 
+with the applications of the theory ranging from relaxed memory models%
+~\cite{Jeffrey-Riely:LICS16, PichonPharabod-Sewell:POPL16, Chakraborty-Vafeiadis:POPL19}
+to model-based mutation testing~\cite{Fellner-al:VMCAI2020}.
+\eupp{perhaps, we could add more citations once the related work is ready.}
+
+The main advantage of the event structures 
+compared to traditional interleaving semantics 
+is that they give a more compact and consice 
+representation of progams' behavior.
+For example, consider the following code snippet.
+
+\input{fig/intro-prog.tex}
+
+Under the interleaving semantics 
+it has $3! = 6$ traces with each trace consisting of $4$ events,
+as depicted on \cref{fig:intro-traces}.
+Event themselves represent atomic side-effects 
+of instructions' execution. In our case 
+it is either a write of a value to a shared variable $\wlab{x}{a}$,
+or a read of a particular value from a shared variable $\rlab{x}{a}$.
+
+\input{fig/intro-traces.tex}
+
+The same information can be encoded in a single 
+event structure containing $6$ events in total
+(see \cref{fig:intro-es}). 
+In the event structure there are two types of edges 
+between the events. The gray arrows $e_1 \arrowCA e_2$ 
+represent the \emph{causality relation}, a 
+partial order reflecting the causal relationship
+between the atomic events of computation.
+The red edges $e_1 \arrowCF e_2$ represent 
+the \emph{conflict relation} which is 
+a symmetric and irreflexive relation 
+encoding mutually exclusive events.
+Each particular trace can be extracted from the event structure
+as a linearisation of some \emph{configuration}, 
+that is a causally-closed and conflict-free 
+subset of events.
+
+\input{fig/intro-es.tex}
+
+\eupp{Continue from here...}
 

### 2021-03-16 papers/SYRCoSE/intro.tex
SYRCoSE: finished with new intro

@@
 
-\eupp{Continue from here...}
+In the programming language and formal semantics research community 
+it become the de-facto standard to complement the theoretical 
+studies with their mechanization in the \emph{proof assistants}
+like Coq, Agda, Isabelle, Arend, and others,
+as this process increases the reliability and reproducibility 
+of the scientific results.
+Yet, as far as we know, there is a little work on 
+mechanization of event structures theory.
+
+Our work is aimed to close this gap. 
+Our goal is to develop a Coq library containing 
+a comprehensive set of common definitions, lemmas, 
+and tactics that would allow the researchers 
+to utilize the theory of event structures 
+for the needs of their domain.
+
+In this work in progress report we sketch 
+the common design princples behind our library
+and give a concrete example of its usage  
+by developing a formal mechanized semantics of simple 
+register machine with shared memory.
 

### 2021-03-16 papers/SYRCoSE/intro.tex
SYRCoSE: minor

@@
 In the programming language and formal semantics research community 
-it become the de-facto standard to complement the theoretical 
+it becomes the de-facto standard to complement the theoretical 
 studies with their mechanization in the \emph{proof assistants}
@@
 register machine with shared memory.
-
-\subsection{Motivation}
-
-% Терия структур событий (СС) была придумана Винскелем в 19??
-году [статья].
-С тех пор она активно развивалась, в различных статьях рассматривалось как различные подходы к этой теории: в [этих статьях] рассмотриваласть категорная семантика, в [этих статьях] [какая-то другая], так и применение СС можно в различных проложениях.
-
-% Для каких же приложений можно использовать СС теорию?
-Например, для задания формальной семантики конкуретных систем [ссылки].
-Истина познается в сравнении, так что давайте попробуем понять неформальный смысл СС на примере задания семнатики какой-нибудь параллельной программы
-
-% Рассмотрим следущую параллельную программу
-\begin{tikzpicture}
-
-      \node (iwy) %[left of = mid, node distance = 1.1cm] 
-      at (-1.3, 2.5)
-      {$x:=1;$};
-
-      \node (irx) [above = -0.5cm of iwy.west, anchor = west] 
-      {$r:=x;$};
-
-      \node (iwx) %[right of = mid, node distance = 0.8cm] 
-       at (0.5, 2.5)
-      {$x:=2;$};
-
-      \draw ($ (iwy.south east) + (0.1,-0.5) $) -- ($ (iwy.south east) + (0.1,0.6) $);
-      \draw ($ (iwx.south west) + (-0.15,-0.5) $) -- ($ (iwx.south west) - (0.15,-0.6) $);
-\end{tikzpicture}
-% Программа разделяется на два потока: в первом мы пишем в х 1, потом читаем из него что-то в r, а во втором пишем в х 2. Вопрос: что окажется в переменной r после исполнения программы?
-% Если мы, например, выполним сначала инструкции из первого потока, а потом из второго то в r окажется 1. Такое поведение отражается на следующей диаграмме:
-
-\begin{tikzpicture}
-    \node (ts1) at (-3,2) {ThreadStart};
-    \node (ts2) at (-1, 2) {ThreadStart};
-    \node (wxo) at (-3,0.2) {$\wlab{x}{1}$};
-    \node (wxt) at (-1,0.2) {$\wlab{x}{2}$};
-    \node (rxo) at (-3, -2.2) {$\rlab{x}{1}$};
-    \draw[ca] (wxo)  -- node{$\leqslant\qquad$}  (ts1);
-    \draw[ca] (wxt)  -- node{$\leqslant\qquad$}  (ts2);
-    \draw[ca] (rxo)  -- node{$\leqslant\qquad$}  (wxo);
-\end{tikzpicture}
-
-% В этой диаграмме мы следим за тем, что происходит с переменной x. Тут есть несколько типов событий:
-% - W(x,u) означает что мы записали в x u
-% - R(x,u) означает что мы прочитали из x u
-% - ThreadStart означает начало потока
-% Так же на этой диаграмме есть отношение $\leq$ -- причинно следственная связь между собыитиями: например оно связывает W(x,1) и R(x,1), что логично, так как запись в x 1 повлекла чтение из x 1 (формально в 2A)
-% Но возможно и другое исполнение программы в котором из x мы прочитаем 2. Ему соотвесвует другая диаграмма:
-\begin{tikzpicture}
-    \node (ts1) at (-3,2) {ThreadStart};
-    \node (ts2) at (-1, 2) {ThreadStart};
-    \node (wxo) at (-3,0.2) {$\wlab{x}{1}$};
-    \node (wxt) at (-1,0.2) {$\wlab{x}{2}$};
-    \node (rxt) at (-3, -2.2) {$\rlab{x}{2}$};
-    \draw[ca] (wxo)  -- node{$\leqslant\qquad$}  (ts1);
-    \draw[ca] (wxt)  -- node{$\leqslant\qquad$}  (ts2);
-    \draw[ca] (rxt)  -- node{$\leqslant\qquad$}  (wxo);
-    \draw[ca] (rxt)  -- node{$\leqslant\qquad$}  (wxt);
-\end{tikzpicture}
-% Заметим что, то тут образовалась новая стрелка из W(x,2) в R(x,2), так как чтение из x 2 не могло бы произойти еслиб мы туда 2 не записывали.
-Однако стрелка из W(x,1) тоже не исчезла, потому что, $x := 1$ в программе стоит перед $r := x$, а значит мы не можем выполнить $r := x$ не выполнив перед этим $x := 1$.
-%  Тут мы видим, что при чтении их x мы столкнулись с некоторой недетерменнированностью, которая отразилось в том, что у нас появилось целое множество графоф, соотвествующих разным исполнения программы
-% По множеству всех таких возможных графов можно понять все возможные поведения программы, однако с ростом недетеменированности в каком-то одном месте растет количество графов, которые отличаются друг от друга только там.
-Хочется использовать такой формализм, который позволил бы описывать локальный недетерменизм более компактно.
-% Здесь нам как раз может помочь второе отношение, присутсвующее в структурах собыий — это отношение конфликта между событиями: e1 # e2. 
-% Оно обозначает, что e1 и e2 не могут произойти в одном исполнении программы например R(x,2) и R(x,1) из примера выше.
-За счет этого в одной СС можно закодировать сразу несколько иполнений программы [статьи], а так же моделировать исполнения, в которых, в ходе оптимизаций, произошли перестановки различных интрукций [про LB и es].
-% Вот так кодируется наш пример с использованием отношкния конфликта
-\begin{tikzpicture}
-    \node (ts1) at (-3,2) {ThreadStart};
-    \node (ts2) at (-1, 2) {ThreadStart};
-    \node (wxo) at (-3,0.2) {$\wlab{x}{1}$};
-    \node (wxt) at (-1,0.2) {$\wlab{x}{2}$};
-    \node (rxt) at (-2, -2.2) {$\rlab{x}{2}$};
-    \node (rxo) at (-4, -2.2) {$\rlab{x}{1}$};
-    \draw[ca] (wxo)  -- node{$\leqslant\qquad$}  (ts1);
-    \draw[ca] (wxt)  -- node{$\leqslant\qquad$}  (ts2);
-    \draw[ca] (rxt)  -- node{$\leqslant\qquad$}  (wxo);
-    \draw[ca] (rxt)  -- node{$\leqslant\qquad$}  (wxt);
-    \draw[ca] (rxo)  -- node{$\leqslant\qquad$}  (wxo);
-    \draw[cf] (rxo)  -- node[above]{$\#$}  (rxt);
-\end{tikzpicture}
-
-\subsection{Problem Statement}
-
-% В существующих работах по структурам событий обычно результаты либо не формализованы ни в каких пруф-ассистантах [ссылки], либо слишком привязаны к конкретной предметной области или задаче [ссылки], что затрудняет переиспользование теории структур событий в других проектах [расписать про это в Related Works].
-
-%Нашей задачей стало решить две, описанные выше, проблемы \\ 1) Формализовать теорию СС в каком-нибудь proof-assistance: предоставить леммы и тактики упрощаюшие работу с СС и доказать основные результаты, описанные в статьях
-
-% \\ 2) Показать, как нашу библиотеку можно использовать для решения прикладных задач, например для задания  семантики абстрактной регистровой машины с слабо-согласованной моделью разделяемой памяти.
-
-%\subsection{Outline} %описание статьи

### 2021-03-18 papers/SYRCoSE/encode_evstr.tex
fix: 'How we encode ...' section moved to separate file

\section{Implementation}

\subsection{How we encode Event Structures}
% Как мы видели ранее в Step-by-step example наша стуркура событий содержит такие функции как frf fpred. Сами события это элементы некоторого множества E (множества итендификаторов), которое должно удволетворять следущим свойствам:
\begin{definition}[identifier set]
  $E$ is an identifier set if:
  \begin{itemize}
    \item There is some $ident_0 \in E$ ~--- first identifier
    \item $E$ has well-founded partitial order $\leqslant_E$
    \item There is some function, that takes some identifier $e$ and returns next fresh identifier. Formally: $fresh : E \to E$, satisfies $$\forall e \in E, \ e <_E fresh(e)$$
  \end{itemize}
\end{definition}
% Так же у нас есть функция которая каждому итендификатору сопоставляет его метку
  $$\ffun{lab} : E \rightharpoonup \lb $$
% где label -- множество всех меток (таких как W(x, 5), R(y, 6), ThreadStart, ThreadEnd etc)
% Мы будем говорить что е это "read"-event c локацией х и значением i если lab(e) = R(x,i), аналогично "write"-event
% В итоге формально мы строим четверку
\begin{definition}[Execution Event Structure]
  $$exec\_event\_structure \ \triangleq \ \langle E, \ \ffun{lab}, \ \fpred, \ \frf \rangle$$
  where
  \begin{itemize}
    \item $E$ ~--- identifiers set
    \item $\fpred$ ~--- function that returns a predsessor (w.r.t. the program order) of event
    \item $\frf$ ~--- function that takes "read"-event $e$ and returns a "write"-event from witch $e$ reads
    \item $\ffun{lab}$ ~--- function that returns event's label
  \end{itemize}
  Also we require
  \begin{itemize}
    \item $\forall e \in E, \ \fpred(e) \leqslant_E e$ and $\frf(e) \leqslant_E e$
    \item $\forall e_1 \ e_2 \in E, \ e_2 = \frf(e_1) \ \Leftrightarrow \ e_1$ is a "read"-event and $e_2$ is a "write"-event with the same location and value
  \end{itemize}
\end{definition}

% Далее для простоты будем отождествлять иденфикаторы с их метками, а Execution Event Structure с их множеством итендификаторов

% Теперь наша задача: определить в этих терминах простые структуры событий. Для этого надо определить частичный порядок <= и отношение конфликта #
% Начнем с отношения causality. В нашей статье мы будем считать, что событие e1 влечет событие e2 если 
\begin{itemize}
  \item Programm instruction, that corresponds to event $e_1$ states before the programm instruction, that corresponds to $e_2$
  \item Or $\ e_2$ is the "read"-event and it reads from "write"-event $e_1$
\end{itemize}
% Или формально:
\begin{itemize}
  \item $\fpred(e_1) \ =\ e_2$
  \item Or $\ \frf(e_1) \ =\ e_2$
\end{itemize}
% Любую функцию f можно считать отношением
  $$ a \ f \ b \ \triangleq\ f(a) \ = \ b$$
% Имея ввиду обозначение выше определим
\begin{definition}[causality relation]
  $$\leqslant \ \triangleq \ (\fpred \cup \frf)^* $$
\end{definition}
% Заметим, что в множестве E уже есть тотальный порядок $\leqslant_E$. Нетрудно видеть что полученнле отношение casuality является его предпорядком [смотри Well-foundedness and monads]
\begin{lemma}
  If $\leqslant$ is a casuality order of $E$ then $\leqslant \ \subset \ \leqslant_E$
\end{lemma}
% Теперь давайте определим отношение конфликта. Посмотрев на отношение конфликта в определении СС [ссылка], мы можем установить следущий факт
\begin{lemma}\label{lma_confl}
  If $\ e_1 \leqslant e_2, \ e_3 \leqslant e_4$ and $e_1 \# e_3$ holds then $e_2 \# e_4$
\end{lemma} 
% То есть отношение конфликта должно "протаскиваться" через отношение <=. Наша интуиция будет заключаться в том, что мы определим "минимальное" отношение конфликта, замкнув, котрое по свойству консистентности [ссылка] мы получим полное отношение #. 
% И так откуда берутся конфликты? [тут должен был заход через Step-by-step example]. 
% И так мы видим, что если у событий есть общий предок (относительно отношения fpred) то эти события происходят в двух разных выполнениях программы и, соответвенно находятся в отношении конфликта
\begin{definition}[immediate conflict relation]
  $$ e_1 \, \ffun{icf} \, e_2 \ \triangleq\ \exists e, \, e = \fpred(e_1) = \fpred(e_2)  $$
\end{definition}
% Далее можно определить отношение конфликта как консистентное замыкание icf
\begin{definition}[conflict relation]
  $$ e_1 \, \# \, e_2 \ \triangleq\ \exists e_3\, e_4, e_3 \, \ffun{icf} \, e_4 \ and \ e_3 \leqslant e_1, \, e_4 \leqslant e_2 $$
\end{definition}
% Однако с этим определением возникает одна проблема. Понятно что отношение определенное таким образом получается симметричным и консистентным. Будет ли оно иррефлексивность? Оказывается, что в общем случае нет. Рассмотрим следущую программу:
\input{fig/encode_prog.tex}
% Давайте как в Step-by-step example построим по шагам фрагмент соответвующей этой программе exec_event_structure до чтения из y
\input{fig/encode_part.tex}
% Видно что наша программа в первом потоку разделилась на два независимых выполнения: выполнение в котором мы прочитали из x 1 (обозначим за выболнение (а)) и выполненние в котором прочитали из х 2 (выполнение (b))
%Теперь мы хотим добавить в нашу структуру событий чтение в y. посмотрим что будет если прочитать при выполнении (а) в у 2 (то есть запись из (b))
\input{fig/encode_cf.tex}
We have $\rlab{x}{1} \leqslant \rlab{y}{2}$, $ \rlab{x}{2} \leqslant \rlab{y}{2}$ and $\rlab{x}{1} \ \# \ \rlab{x}{2}$. Thus $\rlab{y}{2} \ \# \ \rlab{y}{2}$ holds (see~\cref{lma_confl})
% Но отношение конфликта должно быть иррефлексивным. Заметим, однако, что (a) и (b) не могут произойти при одном исполнении программы, а значит и чтения R(y,2) из W(y,2) тут быть не должно, полная структура событий тут должна выглядеть так:
\input{fig/encode_full.tex} \\
% Будем говорить, что сткуктура собырий консистентна если там нет чтений из конфликтующийх с ними записей. Формально
\begin{definition}[consistent Execution Event Structure]
  $$consitent\langle E, \ffun{lab}, \fpred, \frf \rangle \ \triangleq \ \forall e \in E, \ \neg (e \ \# \ \frf e)$$
\end{definition}
% Оказывается этого достатчно чтоб отношение конфликта было иррефлексивным
\begin{lemma}
  Let $E$ be a consistent Execution Event Structure (denote as $cexec\_event\_structure$). Then it's conflict relation $\#$ is irreflexive
\end{lemma}
% Получается что из консистентной exevstr можно получить ПCC

### 2021-03-18 papers/SYRCoSE/intro_to_PES.tex
SYRCoSE: move intro to event structures to separate file

\subsection{Gentle introduction to Event Structures}

Generally, event structures can model a process of computation. 
There are many modifications of events structures from elementary to general, 
and we are using prime event structures, since it is enough for us to model 
concurrent programs. More precisely, they represent that process by a set of events 
and relations between them: 
\begin{definition}
A prime event structure (PES) is a triple $(E, \leq, \#)$, where
\begin{itemize}
  \item E is a set of events
  \item $\leq$ is a causality relation on E such that 
  \begin{itemize}
    \item $ (E, \leq) $ is a partial order
    \item $\lceil e \rceil := \{ e' : e' < e \}$ is finite for every $e \in E$, 
    i.e. every event is caused by a finite set of events
  \end{itemize}
  \item $\#$ is a conflict relation on $E$ such that
  \begin{itemize}
    \item $\#$ is irreflexive and symmetric
    \item it satisfies a consistency predicate:
    $$\forall e_1, e_2, e_3 \in E \ e_1 \# e_2 \land e_2 \leq e_3 \Rightarrow e_1 \# e_3.$$
  \end{itemize}
\end{itemize}
\end{definition}

Conflict relation is used to model conflicting events like two \texttt{write}'s 
of different values to the same location.
Consistency ensures that if some events are in conflict, then events 
following these events are also in conflict.

\begin{definition}
A configuration of PES $(E, \leq, \#)$ is a subset $X \subseteq E$ such that
\begin{itemize}
  \item it is conflict-free
  \item it is left-closed, i.e. $$\forall e_1, e_2 \in E \ e_2 \in X \land e_1 \leq e_2 \Rightarrow e_1 \in E.$$
\end{itemize}
\end{definition}

Configurations are used to model a history of computation up to a certain 
execution point.

We formally proved the following simple facts about prime event structures:

\begin{lemma}
  A set $\lceil e \rceil$ is conflict-free for any $e \in E$.
\end{lemma}

And a more general one:

\begin{lemma}
  A set $\lceil e \rceil$ is a configuration for any $e \in E$.
\end{lemma}

### 2021-03-18 papers/SYRCoSE/regmachine.tex

\subsection{Register Machine and thread semantics}
Now we are going to define language of simple register machine with shared memory.
\newcommand{\ip}{\ffun{ip}}
\newcommand{\regmap}{\ffun{regmap}}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\[
\begin{array}{rcll}
  \texttt{P} &::=& \texttt{ST}^* \\
  \texttt{ST} & ::= & \texttt{I}^* \\
  \texttt{I} & ::= & \texttt{Reg := Val} \\
                 & | & \texttt{Loc := Val} \\
                 & | & \texttt{Reg := Loc} \\
                 & | & \texttt{CJmp Reg $\mathbb{N}$} \\
  \texttt{Reg}            & ::= & r \, | \, r' \, | ... \\ 
  \texttt{Loc}            & ::= & x \, | \,y \,| ... \\ 
\end{array}
\] where
\begin{itemize}
  \item \texttt{P} -- program is a sequence of single thread progams
  \item \texttt{ST} -- single thread program is a sequence of instructions
  \item \texttt{I} -- instruction can be either write to some register of some value or location value, either write to some location or conditional jump
\end{itemize}
Here \texttt{Reg} are registers (denoted as $r,r',...$) ~--- variables that should be different for different single thread progams. In contrast \texttt{Loc} are local variables (denoted as $x,y,...$) ~--- this is our shared (between different threads) memory. Also \texttt{Val} is the type of values (denoted as $i,j,...$) in our program, for simplicity let it be $\mathbb{N}$ or $\mathbb{Z}$. \\
We are interested at looking after shared memory, so in our event structures we have events that represent actions only at local variables. \\
Next we define semantics of single thread program. Let's consider some single-thread program (\texttt{ST} in our notation) and define thread state:
\begin{definition}[Thread State]
  $$thread\_state \ \triangleq \ \langle \ffun{ip}, \ffun{regmap} \rangle,$$ where
  \begin{itemize}
    \item $\ffun{ip}$ ~--- our place in program (ie. line number) 
    \item $\ffun{regmap}$ ~--- map from registers to values
  \end{itemize}
\end{definition}
If we have some $p : \texttt{ST}$, and some line number $\ffun{ip}$ we will denote by $p[\ffun{ip}]$ -- $\ffun{ip}$'s instuction of $p$. \\
Then we define one-step semantics relation $st, p \xRightarrow{l} cont\_st$, where
\begin{itemize}
  \item $st$ ~--- thread state
  \item $p$ ~--- our single-thead programm
  \item $l$ ~--- label that denotes action on shared memory (can be with hole)
  \item $cont\_st$ ~--- continuation of thread state (ie. function from \texttt{Val} to $thread\_state$). We clarify why we need continuation here a bit later
\end{itemize}
If we have some thread state $st$, we can identify $\lambda\, x,\, st$ (ie. constant continuation) with $st$ itself. Also by $f[r \leftarrow i]$ we will denote function that sends $r$ to $i$ and $r'$ to $f(r')$, for all $r' \neq r$

\input{fig/thread_sem.tex}

Every rule in \cref{fig:thread_sem} should be clear (see [TODO]), except the last one. Consider we want to read some local variable $x$ in some thread. Another threads can write something to this $x$ too, and it is possible to read from some of these writes. While building semantics of our whole programm we will have enough information to decide from wich write we can read. But for now let us consider that our side-effect is the label with the hole, and our result will be a continuation of a thread state (ie. some function $cont\_st : \texttt{Val} \to thread\_state$).

### 2021-03-18 papers/SYRCoSE/encode_evstr.tex
feat: add part of small-step semantics section

@@
 \begin{definition}[identifier set]
-  $E$ is an identifier set if:
+  $\mathcal{E}$ is an identifier set if:
   \begin{itemize}
-    \item There is some $ident_0 \in E$ ~--- first identifier
-    \item $E$ has well-founded partitial order $\leqslant_E$
-    \item There is some function, that takes some identifier $e$ and returns next fresh identifier. Formally: $fresh : E \to E$, satisfies $$\forall e \in E, \ e <_E fresh(e)$$
+    \item There is some $ident_0 \in \mathcal{E}$ ~--- first identifier
+    \item $\mathcal{E}$ has well-founded partitial order $\leqslant_\mathcal{E}$
+    \item There is some function, that takes some finite identifies set $E \subset \mathcal{E}$ and returns a fresh identifier.
+Formally: $fresh : \mathcal{P}_{fin}(\mathcal{E}) \to \mathcal{E}$, satisfies $$\forall e \in E, \ e <_\mathcal{E} fresh(E)$$
   \end{itemize}
 \end{definition}
-% Так же у нас есть функция которая каждому итендификатору сопоставляет его метку
+% Пусть E это множества всех событий программы тогда у нас есть
   $$\ffun{lab} : E \rightharpoonup \lb $$
@@
   \begin{itemize}
-    \item $E$ ~--- identifiers set
+    \item $E$ ~--- finite subset of some identifiers set $\mathcal{E}$
     \item $\fpred$ ~--- function that returns a predsessor (w.r.t. the program order) of event
     \item $\frf$ ~--- function that takes "read"-event $e$ and returns a "write"-event from witch $e$ reads
-    \item $\ffun{lab}$ ~--- function that returns event's label
+    \item $\ffun{lab} : E \to \lb$ ~--- function that returns event's label
   \end{itemize}

### 2021-03-18 papers/SYRCoSE/regmachine.tex
feat: add part of small-step semantics section

@@
 Now we are going to define language of simple register machine with shared memory.
-\newcommand{\ip}{\ffun{ip}}
-\newcommand{\regmap}{\ffun{regmap}}
-\newcommand{\la}{\langle}
-\newcommand{\ra}{\rangle}
 \[
@@
                  & | & \texttt{Reg := Loc} \\
-                 & | & \texttt{CJmp Reg $\mathbb{N}$} \\
+                 & | & \texttt{CJmp Reg $\mathbb{N} \backslash \{0\}$} \\
   \texttt{Reg}            & ::= & r \, | \, r' \, | ... \\ 
@@
 \begin{definition}[Thread State]
-  $$thread\_state \ \triangleq \ \langle \ffun{ip}, \ffun{regmap} \rangle,$$ where
+  $$thread\_state \ \triangleq \ \langle \ffun{ip}, \regmap \rangle,$$ where
   \begin{itemize}
     \item $\ffun{ip}$ ~--- our place in program (ie. line number) 
-    \item $\ffun{regmap}$ ~--- map from registers to values
+    \item $\regmap$ ~--- map from registers to values
   \end{itemize}
 \end{definition}
-If we have some $p : \texttt{ST}$, and some line number $\ffun{ip}$ we will denote by $p[\ffun{ip}]$ -- $\ffun{ip}$'s instuction of $p$. \\
+If we have some $p : \texttt{ST}$, and some line number $\ffun{ip}$ we will denote by $p[\ffun{ip}]$ -- $\ffun{ip}$'s instuction of $p$ (we start numeration from 1).  \\
 Then we define one-step semantics relation $st, p \xRightarrow{l} cont\_st$, where

### 2021-03-18 papers/SYRCoSE/small_step.tex
feat: add part of small-step semantics section

\subsection{Construction small-step semantics}
Consider we have some Execution Event Structure $es = \la E, \ffun{lab}, \fpred, \frf \ra$, $pr \in E$ ~--- the $es$'s event and some fresh identifier $e = fresh(E)$. Let's define expansion $es$ with the event $e$ depending on its label:
\begin{definition}[Add event]
  \begin{itemize}
    \item $es + \la \wlab{x}{i}, \ pr \ra \ \triangleq \vspace{1mm} \\ \ 
    \la E \cup \{e\}, \
        \ffun{lab}[e \leftarrow \wlab{x}{i}], \
        \fpred[e \leftarrow pr], \
        \frf \ra\ $
        \vspace{1mm}
    \item $es + \la \text{ThreadStart} \ra \ \triangleq \vspace{1mm} \\
    \la E \cup \{e\}, \
    \ffun{lab}[e \leftarrow \text{ThreadStart}], \
    \fpred, \
    \frf \ra\ $
    \vspace{1mm}
    \item $es + \la \rlab{x}{i}, \ pr, \ wr \ra \ \triangleq$
    $$\la E \cup \{e\}, \
    \ffun{lab}[e \leftarrow \rlab{x}{i}], \
    \fpred[e \leftarrow pr], \
    \frf[e \leftarrow wr] \ra\ $$
  \end{itemize}
\end{definition}
Then we define programm configuration:
\begin{definition}[configuration]
  $$config \ \triangleq \ \la \es, \tmap \ra$$
  where
  \begin{itemize}
    \item $\es$ ~--- our current consistent Execution Event Structure
    \item $\tmap$ ~--- mapping from events in $\es$ to the pair of corresponding [TODO]
  \end{itemize}
\end{definition}

By $\es.E$, $\es.\ffun{lab}$, $\es.fr$ we denote $E$, $\ffun{lab}$ and $fresh(E)$ respectively, where $\es = \la E, \ffun{lab}, \fpred, \frf \ra$. Also by $\es.tid$ we denote the smallest number of thread that wasn't started yet. \\
If we have some $p : \texttt{P}$, and $t \in \mathbb{N}$ ~--- thread number, $p[t] :=$ $t$'s thread of programm $p$.

Also we will need 'can-read-from relation':
\begin{definition}[Can-read-from relation]
  If $l, l' \in \lb$ we can define
  $$ l \ll l' \ \triangleq \ l = \rlab{x}{i} \ \text{and} \ l' = \wlab{x}{i}$$
\end{definition}

\input{fig/small_step.tex}
[TODO: small step rules explanation]

### 2021-03-24 papers/SYRCoSE/encode_evstr.tex
SYRCoSE: add some text

@@
 \subsection{How we encode Event Structures}
-% Как мы видели ранее в Step-by-step example наша стуркура событий содержит такие функции как frf fpred.
-Сами события это элементы некоторого множества E (множества итендификаторов), которое должно удволетворять следущим свойствам:
+As we saw earlier in Step by Step Example, our event structure contains functions such as $\frf$ and $\fpred$.
+The events themselves are elements of some set $\mathcal{E}$ (set of identifiers), which must satisfy the following properties
 \begin{definition}[identifier set]
@@
     \item $\mathcal{E}$ has well-founded partitial order $\leqslant_\mathcal{E}$
-    \item There is some function, that takes some finite identifies set $E \subset \mathcal{E}$ and returns a fresh identifier.
+    \item There is a function, that takes some finite $E \subset \mathcal{E}$ and returns a fresh identifier.
 Formally: $fresh : \mathcal{P}_{fin}(\mathcal{E}) \to \mathcal{E}$, satisfies $$\forall e \in E, \ e <_\mathcal{E} fresh(E)$$
@@
 \end{definition}
-% Пусть E это множества всех событий программы тогда у нас есть
-  $$\ffun{lab} : E \rightharpoonup \lb $$
-% где label -- множество всех меток (таких как W(x, 5), R(y, 6), ThreadStart, ThreadEnd etc)
-% Мы будем говорить что е это "read"-event c локацией х и значением i если lab(e) = R(x,i), аналогично "write"-event
-% В итоге формально мы строим четверку
+We can fix any $\mathcal{E}$ (for example $\mathbb{N}$) and theat programm's events as a finite subset of $\mathcal{E}$
+
+Let $E$ be the set of all program events, then we have
+  $$\ffun{lab} : E \to \lb $$
+where $\lb$ is the label's set (such that $\wlab{x}{5}, \rlab{y}{6}$ ThreadStart, etc.)
+
+We will say that $e$ is a "read" -event with location x and value i if $\ffun{lab}(e) = \rlab{x}{i}$, similarly to "write" -event
+
+As a result, we formally build a four
 \begin{definition}[Execution Event Structure]
@@
 \end{definition}
+Further, for simplicity, we will identify identifiers with their labels, and Execution Event Structure with their set of identifiers
 
-% Далее для простоты будем отождествлять иденфикаторы с их метками, а Execution Event Structure с их множеством итендификаторов
+Now our goal is to define a prime event structures in these terms. To do this, one need to come up with a causality relation $\leqslant$ and a conflict relation $\#$
 
-% Теперь наша задача: определить в этих терминах простые структуры событий. Для этого надо определить частичный порядок <= и отношение конфликта #
-% Начнем с отношения causality. В нашей статье мы будем считать, что событие e1 влечет событие e2 если 
+Let's start with the causality. In our article, we will assume that the event $e_1$ entails the event $e_2$ if
 \begin{itemize}
@@
 \end{itemize}
-% Или формально:
+Or formally:
 \begin{itemize}
-  \item $\fpred(e_1) \ =\ e_2$
-  \item Or $\ \frf(e_1) \ =\ e_2$
+  \item $\fpred(e_2) \ =\ e_1$
+  \item Or $\ \frf(e_2) \ =\ e_1$
 \end{itemize}
-% Любую функцию f можно считать отношением
-  $$ a \ f \ b \ \triangleq\ f(a) \ = \ b$$
-% Имея ввиду обозначение выше определим
+Any function $f$ can be treated as a relation
+  $$ a \ f \ b \ \triangleq\ f(b) \ = \ a$$
+With the notation above, we define
+
+\begin{center}
 \begin{definition}[causality relation]
@@
 \end{definition}
-% Заметим, что в множестве E уже есть тотальный порядок $\leqslant_E$. Нетрудно видеть что полученнле отношение casuality является его предпорядком [смотри Well-foundedness and monads]
+\end{center}
+
+Note that the set $E \subset \mathcal{E}$ already contains the total order $ \leqslant_\mathcal{E} $.
+One can note that the resulting casuality relation is its suborder [TODO:see Well-foundedness and monads]
+
+\begin{center}
 \begin{lemma}
-  If $\leqslant$ is a casuality order of $E$ then $\leqslant \ \subset \ \leqslant_E$
+  If $\leqslant$ is a casuality order of $E \subset \mathcal{E}$ then $\leqslant \ \subset \ \leqslant_\mathcal{E}$
 \end{lemma}
-% Теперь давайте определим отношение конфликта. Посмотрев на отношение конфликта в определении СС [ссылка], мы можем установить следущий факт
+\end{center}
+
+Now let's define a conflict relation. Looking at the conflict relation in \cref{es_def}, we can establish the following fact:
+\begin{center}
 \begin{lemma}\label{lma_confl}
@@
 \end{lemma} 
-% То есть отношение конфликта должно "протаскиваться" через отношение <=.
-Наша интуиция будет заключаться в том, что мы определим "минимальное" отношение конфликта, замкнув, котрое по свойству консистентности [ссылка] мы получим полное отношение #.
-% И так откуда берутся конфликты? [тут должен был заход через Step-by-step example]. 
-% И так мы видим, что если у событий есть общий предок (относительно отношения fpred) то эти события происходят в двух разных выполнениях программы и, соответвенно находятся в отношении конфликта
+\end{center}
+That is, the conflict relation must be "dragged" through the $\leqslant$ relation.
+Our intuition will be as follows: we, first, define a "minimal" conflict relation, by closing which under the consistency property (\cref{lma_confl}), we finally get a full relation $\#$.
+
+So where do conflicts come from?
+
+From the step-by-step example, we see that if events have a common predsessor (w.r.t. the $\fpred$ relation), then they occur in two different program executions and, accordingly, are conflicting
 \begin{definition}[immediate conflict relation]
@@
 \end{definition}
-% Далее можно определить отношение конфликта как консистентное замыкание icf
+Next, one can define the conflict relation as a consistent closure of $\ffun{icf}$
 \begin{definition}[conflict relation]
@@
 \end{definition}
-% Однако с этим определением возникает одна проблема.
-Понятно что отношение определенное таким образом получается симметричным и консистентным.
-Будет ли оно иррефлексивность?
-Оказывается, что в общем случае нет.
-Рассмотрим следущую программу:
+However, there is one problem with such definition.
+It is clear that the relation defined in this way turns out to be symmetric and consistent.
+Will it be irreflexive?
+It turns out that in the general case, no.
+Consider the following program:
 \input{fig/encode_prog.tex}
-% Давайте как в Step-by-step example построим по шагам фрагмент соответвующей этой программе exec_event_structure до чтения из y
+
+Let's build, as in the Step-by-step example, a fragment of $exec\_event\_structure$ corresponding to this program before reading from $y$
 \input{fig/encode_part.tex}
-% Видно что наша программа в первом потоку разделилась на два независимых выполнения: выполнение в котором мы прочитали из x 1 (обозначим за выболнение (а)) и выполненние в котором прочитали из х 2 (выполнение (b))
-%Теперь мы хотим добавить в нашу структуру событий чтение в y. посмотрим что будет если прочитать при выполнении (а) в у 2 (то есть запись из (b))
+
+One can observe, that our program in the first thread was divided into two independent executions: execution, in which we read from $x$ value 1 (denote by execution (a)) and execution, in which we read from $x$ 2 (execution (b)).
+
+Now we want to add a read-from-$y$ event to our event structure. Let's see what is going to happen if we will add "read"-event, reading from $\wlab{y}{2}$ (see \cref{fig:cf}).
 \input{fig/encode_cf.tex}
+
 We have $\rlab{x}{1} \leqslant \rlab{y}{2}$, $ \rlab{x}{2} \leqslant \rlab{y}{2}$ and $\rlab{x}{1} \ \# \ \rlab{x}{2}$. Thus $\rlab{y}{2} \ \# \ \rlab{y}{2}$ holds (see~\cref{lma_confl})
-% Но отношение конфликта должно быть иррефлексивным.
-Заметим, однако, что (a) и (b) не могут произойти при одном исполнении программы, а значит и чтения R(y,2) из W(y,2) тут быть не должно, полная структура событий тут должна выглядеть так:
-\input{fig/encode_full.tex} \\
-% Будем говорить, что сткуктура собырий консистентна если там нет чтений из конфликтующийх с ними записей. Формально
+
+To deal with this program, we need to impose a constraint on the event structure.
+
+We will say that the event structure is consistent if there are no reads from conflicting writes. Formally:
 \begin{definition}[consistent Execution Event Structure]
@@
 \end{definition}
-% Оказывается этого достатчно чтоб отношение конфликта было иррефлексивным
+It turns out that this is enough for the conflict relation to be irreflexive. (see \cref{fig:full})
 \begin{lemma}
@@
 \end{lemma}
-% Получается что из консистентной exevstr можно получить ПCC
+\input{fig/encode_full.tex} 
+So, we see that from the consistent event structure one can get the Prime Event Structure

### 2021-03-24 papers/SYRCoSE/regmachine.tex
SYRCoSE: add some text

@@
   \item \texttt{ST} -- single thread program is a sequence of instructions
-  \item \texttt{I} -- instruction can be either write to some register of some value or location value, either write to some location or conditional jump
+  \item \texttt{I} -- instruction can be 
+  \begin{itemize}
+    \item \texttt{Reg := Val} -- writing some value to the register
+    \item \texttt{Reg := Loc} -- reading a value from the shared-memory-location and writing it to the register
+    \item \texttt{Loc := Val} -- writing some value to the shared-memory-location
+    \item CJmp Reg $\mathbb{N} \backslash \{0\}$ -- conditional jump
+  \end{itemize}
 \end{itemize}
@@
 We are interested at looking after shared memory, so in our event structures we have events that represent actions only at local variables. \\
-Next we define semantics of single thread program. Let's consider some single-thread program (\texttt{ST} in our notation) and define thread state:
+Next we define the semantics of a single thread program. Let's consider some single-thread program (\texttt{ST} in our notation) and define a thread state:
 \begin{definition}[Thread State]
-  $$thread\_state \ \triangleq \ \langle \ffun{ip}, \regmap \rangle,$$ where
+  $$thread\_state \ \triangleq \ \langle \ip, \regmap \rangle,$$ where
   \begin{itemize}
-    \item $\ffun{ip}$ ~--- our place in program (ie. line number) 
+    \item $\ip$ ~--- our place in program (ie. line number) 
     \item $\regmap$ ~--- map from registers to values
@@
 \end{definition}
-If we have some $p : \texttt{ST}$, and some line number $\ffun{ip}$ we will denote by $p[\ffun{ip}]$ -- $\ffun{ip}$'s instuction of $p$ (we start numeration from 1).  \\
-Then we define one-step semantics relation $st, p \xRightarrow{l} cont\_st$, where
+If we have some $p : \texttt{ST}$, and some line number $\ip$ we will denote by $p[\ip]$ -- $\ip$'s instuction of $p$ (we start numeration from 1).  \\
+Then we define one-step semantics relation $\st, p \xRightarrow{l} \cst$, where
 \begin{itemize}
-  \item $st$ ~--- thread state
+  \item $\st$ ~--- thread state
   \item $p$ ~--- our single-thead programm
   \item $l$ ~--- label that denotes action on shared memory (can be with hole)
-  \item $cont\_st$ ~--- continuation of thread state (ie. function from \texttt{Val} to $thread\_state$). We clarify why we need continuation here a bit later
+  \item $\cst$ ~--- continuation of thread state (ie. function from \texttt{Val} to $thread\_state$). We clarify why we need continuation here a bit later
 \end{itemize}
-If we have some thread state $st$, we can identify $\lambda\, x,\, st$ (ie.
-constant continuation) with $st$ itself.
+If we have some thread state $\st$, we can identify $\lambda\, x,\, \st$ (ie.
+constant continuation) with $\st$ itself.
 Also by $f[r \leftarrow i]$ we will denote function that sends $r$ to $i$ and $r'$ to $f(r')$, for all $r' \neq r$
@@
 
-Every rule in \cref{fig:thread_sem} should be clear (see [TODO]), except the last one.
+Every rule in \cref{fig:thread_sem} should be clear (see [TODO]), except maybe the last one.
 Consider we want to read some local variable $x$ in some thread.
@@
 While building semantics of our whole programm we will have enough information to decide from wich write we can read.
-But for now let us consider that our side-effect is the label with the hole, and our result will be a continuation of a thread state (ie.
-some function $cont\_st : \texttt{Val} \to thread\_state$).
+But for now let us assume that our side-effect is the label with the hole, and our result is a continuation of a thread state (ie.
+some function $\cst : \texttt{Val} \to thread\_state$).

### 2021-03-24 papers/SYRCoSE/small_step.tex
SYRCoSE: add some text

@@
 Consider we have some Execution Event Structure $es = \la E, \ffun{lab}, \fpred, \frf \ra$, $pr \in E$ ~--- the $es$'s event and some fresh identifier $e = fresh(E)$.
-Let's define expansion $es$ with the event $e$ depending on its label:
+Let's define expansion $es$ with $e$ depending on it's label:
 \begin{definition}[Add event]
@@
   \end{itemize}
+\input{fig/small_step.tex}
 \end{definition}
@@
     \item $\es$ ~--- our current consistent Execution Event Structure
-    \item $\tmap$ ~--- mapping from events in $\es$ to the pair of corresponding [TODO]
+    \item $\tmap$ ~--- map that takes event $e$ from $\es$ and returns some pair.
+First component of this pair is a state of some thread after execution of $e$'s memory-side-effect.
+Our programm contains several threads, so we need to specify number of thread (where everything is going on) in the second component
   \end{itemize}
 \end{definition}
+So programm configuration describes current Execution Event Structure, where we map each event to corresponding thread state.
+Thus, to make our semantics step we should select some event and make progress in it's thread state.
+Before putting it formally we will need some notations:
 
@@
 
-\input{fig/small_step.tex}
-[TODO: small step rules explanation]
+One can find definition of small-step semantics in \cref{fig:smallstep} \\
+First look at $\to$ThreadStart rule. In this rule we are starting a new thread by adding some fresh identifier to $\es.E$ with label -- ThreadStart. \\
+If we proceeded in thread state of some event $e$ without any memory-side-effect, we should just update $e$'s thread state (see $\to\_$). \\
+Another possibility of memory-side-effect is write to some shared-memory location.
+In this case we should just add a new write-event to our Execution Event Structure, and map it to the thread state after our side-effect (see $\to\wlab{x}{i}$).
+\\
+We observe the most interesting situation when our side-effect was $\rlab{x}{\_}$ (see $\to\rlab{x}{i}$).
+That means that we wanted to read from some shared-memory location.
+We need to choose some write-event $w \in \es.E$ and read value from it.
+It's clear that $w$ should have label $\wlab{x}{i}$, for some $i$.
+\\
+Unfortunately, in the last step we could read from conflicting event $w$.
+So after making $\to$ step, it is necessary to chek if our final $\es'$ is consistent (see small-step rule).
+After all, we obtain one-step relation $\Rightarrow$.

### 2021-03-31 papers/SYRCoSE/regmachine.tex
SYRCoSE: WIP --- reworking text

--- 2021-03-18 papers/SYRCoSE/regmachine.tex
@@
-\subsection{Register Machine and thread semantics}
-Now we are going to define language of simple register machine with shared memory.
+\subsection{Register Machine}
+
+For our case study we use a simple idealized model 
+of a register machine, which consists of a finite 
+sequence of instructions, an instruction pointer 
+and an infinite set of registers. 
+The syntax of the machine's language is 
+shown on~\cref{fig:regmachine-syntax}.
+
+\begin{figure}[h!]
 \[
-\begin{array}{rcll}
-  \texttt{P} &::=& \texttt{ST}^* \\
-  \texttt{ST} & ::= & \texttt{I}^* \\
-  \texttt{I} & ::= & \texttt{Reg := Val} \\
-                 & | & \texttt{Loc := Val} \\
-                 & | & \texttt{Reg := Loc} \\
-                 & | & \texttt{CJmp Reg $\mathbb{N} \backslash \{0\}$} \\
-  \texttt{Reg}            & ::= & r \, | \, r' \, | ... \\ 
-  \texttt{Loc}            & ::= & x \, | \,y \,| ... \\ 
+\begin{array}{rclclll}
+  P &\in& \Prog   &::=& i_1 \seq \dots \seq i_n           & \text{program}                 \\
+
+  I &\in& \Inst     &::=&                                 & \text{instruction}             \\
+              & &   & | & \assignInst{r}{v}               & \text{assign to register}      \\
+              & &   & | & \opInst{\otimes}{r_1}{r_2}{r_3}  & \text{apply binary operation}  \\
+              & &   & | & \jumpInst{r}{\ip{i}}            & \text{conditional jump}        \\
+              & &   & | & \readInst{r}{x}                 & \text{read from memory}        \\
+              & &   & | & \writeInst{x}{v}                & \text{write to memory}         \\
+
+  r      &\in& \Reg &   &                                & \text{thread-local register}   \\ 
+  x      &\in& \Loc &   &                                & \text{shared memory location}  \\ 
+  v      &\in& \Z   &   &                                & \text{value}                   \\
+  \ip{i} &\in& \N   &   &                                & \text{instruction label}       \\
+
 \end{array}
-\] where
-\begin{itemize}
-  \item \texttt{P} -- program is a sequence of single thread progams
-  \item \texttt{ST} -- single thread program is a sequence of instructions
-  \item \texttt{I} -- instruction can be either write to some register of some value or location value, either write to some location or conditional jump
-\end{itemize}
-Here \texttt{Reg} are registers (denoted as $r,r',...$) ~--- variables that should be different for different single thread progams.
-In contrast \texttt{Loc} are local variables (denoted as $x,y,...$) ~--- this is our shared (between different threads) memory.
-Also \texttt{Val} is the type of values (denoted as $i,j,...$) in our program, for simplicity let it be $\mathbb{N}$ or $\mathbb{Z}$.
-\\
-We are interested at looking after shared memory, so in our event structures we have events that represent actions only at local variables. \\
-Next we define semantics of single thread program. Let's consider some single-thread program (\texttt{ST} in our notation) and define thread state:
-\begin{definition}[Thread State]
-  $$thread\_state \ \triangleq \ \langle \ffun{ip}, \regmap \rangle,$$ where
-  \begin{itemize}
-    \item $\ffun{ip}$ ~--- our place in program (ie. line number) 
-    \item $\regmap$ ~--- map from registers to values
-  \end{itemize}
-\end{definition}
-If we have some $p : \texttt{ST}$, and some line number $\ffun{ip}$ we will denote by $p[\ffun{ip}]$ -- $\ffun{ip}$'s instuction of $p$ (we start numeration from 1).  \\
-Then we define one-step semantics relation $st, p \xRightarrow{l} cont\_st$, where
-\begin{itemize}
-  \item $st$ ~--- thread state
-  \item $p$ ~--- our single-thead programm
-  \item $l$ ~--- label that denotes action on shared memory (can be with hole)
-  \item $cont\_st$ ~--- continuation of thread state (ie. function from \texttt{Val} to $thread\_state$). We clarify why we need continuation here a bit later
-\end{itemize}
-If we have some thread state $st$, we can identify $\lambda\, x,\, st$ (ie.
-constant continuation) with $st$ itself.
-Also by $f[r \leftarrow i]$ we will denote function that sends $r$ to $i$ and $r'$ to $f(r')$, for all $r' \neq r$
+\] 
+\caption{Syntax of the register machine}
+\label{fig:regmachine-syntax}
+\end{figure}
 
-\input{fig/thread_sem.tex}
+We first present the semantics of a single thread program.
+In this semantics memory access instructions 
+does not actually operate on shared memory,
+but rather produce a \emph{label}
+denoting the side-effect of the operation 
+(see \cref{fig:label-syntax}).
+This encoding allows us to decouple 
+the semantics of the register machine 
+from a \emph{memory model}~\cite{\todo{}}.
 
-Every rule in \cref{fig:thread_sem} should be clear (see [TODO]), except the last one.
-Consider we want to read some local variable $x$ in some thread.
-Another threads can write something to this $x$ too, and it is possible to read from some of these writes.
-While building semantics of our whole programm we will have enough information to decide from wich write we can read.
-But for now let us consider that our side-effect is the label with the hole, and our result will be a continuation of a thread state (ie.
-some function $cont\_st : \texttt{Val} \to thread\_state$).
+\begin{figure}[h!]
+\[
+\begin{array}{rclclll}
+
+  \ell &\in& \Lab  &::=&                                & \text{}             \\
+              & &  & | & \rlab{x}{v}                    & \text{read of value $v$ from location $x$}   \\
+              & &  & | & \wlab{x}{v}                    & \text{write of value $v$ to location $x$}    \\
+
+\end{array}
+\] 
+\caption{Syntax of Labels}
+\label{fig:label-syntax}
+\end{figure}
+
+The semantics is given in a form of 
+\emph{labeled transition system}~\cite{\todo{}}.
+State of the system consists of instruction pointer
+and a map from registers to their values, 
+as shown in \cref{fig:regmachine-thrdstate}. 
+The rules of the semantics are standard 
+(see \cref{fig:regmachine-thrdsem}).
+
+\begin{figure}[h!]
+\[
+\begin{array}{rclclll}
+
+  s          &\in& \ThrdState    &::=& \tup{\ip{i}, \regmap} &                         \\
+  \ip{i}     &\in& \N            &   &                    & \text{instruction pointer} \\
+  \regmap    &\in& \Reg \fun \Z  &   &                    & \text{registers mapping}   \\
+
+\end{array}
+\] 
+\caption{Thread state of register machine}
+\label{fig:regmachine-thrdstate}
+\end{figure}
+
+\input{fig/regmachine-thrdsem.tex}

### 2021-04-01 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: WIP --- reworking the text

--- 2021-03-24 papers/SYRCoSE/encode_evstr.tex
@@
-\section{Implementation}
+\subsection{Event Structure of Register Machine}
 
-\subsection{How we encode Event Structures}
+In this section we present alternative operational semantics
+which constructs a prime event structure encoding a 
+behavior of the register machine.
+
+The event structure is constructed incrementally 
+in a step-by-step fashion by adding a single event 
+on each step. In order to generate a new event 
+on each step we require that events behave as \emph{identifiers}.
+
+\begin{definition}[Identifiers set]
+  We say that set $\eventSet$ together with strict partial order $\identOrd$
+  form an identifier set if:
+  \begin{itemize}
+    \item there exist distinguished initial identifier $e_0 \in \eventSet$;
+    \item there exist function $\fresh : \eventSet \fun \eventSet$ which 
+      generates a new fresh identifier, \sth
+      $$ \forall e \in \eventSet \ldotp e \identOrd \fresh(e) $$
+  \end{itemize}
+\end{definition}
+
+Following the theory of axiomatic weak memory models~\cite{\todo{}},
+we define the causality relation of the register machine's event structure
+as a transitive closure of two relations~---~\emph{program order} 
+and \emph{reads-from}, denoted as $\lPO$ and $\lRF$ correspondingly.
+
+$$ \caOrd \ \defeq \ (\lPO \cup \lRF)^* $$
+
+Program order tracks precedence of events within a single thread. 
+Reads-from relation captures the flow of values from 
+write events to read events, and ensures that values 
+do not appear out of thin-air~\cite{\todo{}}.
+
+In order to construct $\lPO$ and $\lRF$ incrementally
+we represent them via their inverse covering functions.
+
+\begin{definition}[Covering]
+  Let $\leqslant$ be a partial order. 
+  Then $lessdot$ is covering relation \wrt $\leqslant$ whenever 
+  $x \lessdot y$ is true if and only if $x < y$ and 
+  there is no $z$ \sth $x < z$ and $z < y$.
+  A (non-deterministic) function $f : A \fun \pwset{A}$ is a covering function if
+  its corresponding relation, \ie ${\frel{f} \defeq \set{\tup{x, y} ~|~ y \in f(x)}}$, 
+  is a covering relation.
+\end{definition}
+
+We use the inverse covering function because it is 
+more convinient in our setting. Indeed, the semanics adds new event
+at each step. Then it is convinient to require that in addition 
+the small-step relation is provided with the $\lPO$ and $\lRF$
+predecessors of a new event.
+
+\[\def\arraystretch{1.5}
+\begin{array}{rcl|rcl}
+
+  \lessdot_{\lPO} &\defeq& {\frel{f_{\lPO}}}^{-1}  & \lessdot_{\lRF} &\defeq& {\frel{f_{\lRF}}}^{-1} \\
+  \lPO           &\defeq& \lessdot_{\lPO}^+      & \lRF           &\defeq& \lessdot_{\lRF}       \\
+
+\end{array}
+\]
+
+The primitive conflict relation is generated by the $f_{\lPO}$ function.
+The two events are considered to be in primitive conflict if they are
+not equal and have a commond $\lPO$ predecessor. 
+For this definition to work properly, we also need 
+to assume that in the event structure each thread has 
+a special initial event labeled by a distinguished 
+\emph{thread start} label $\tslab{}$.
+
+$$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge f_{\lPO}(e_1) = f_{\lPO}(e2) $$
+
+\eupp{Continue from here}
+
 As we saw earlier in Step by Step Example, our event structure contains functions such as $\frf$ and $\fpred$.
 The events themselves are elements of some set $\mathcal{E}$ (set of identifiers), which must satisfy the following properties
-\begin{definition}[identifier set]
-  $\mathcal{E}$ is an identifier set if:
-  \begin{itemize}
-    \item There is some $ident_0 \in \mathcal{E}$ ~--- first identifier
-    \item $\mathcal{E}$ has well-founded partitial order $\leqslant_\mathcal{E}$
-    \item There is a function, that takes some finite $E \subset \mathcal{E}$ and returns a fresh identifier.
-Formally: $fresh : \mathcal{P}_{fin}(\mathcal{E}) \to \mathcal{E}$, satisfies $$\forall e \in E, \ e <_\mathcal{E} fresh(E)$$
-  \end{itemize}
-\end{definition}
 We can fix any $\mathcal{E}$ (for example $\mathbb{N}$) and theat programm's events as a finite subset of $\mathcal{E}$
@@
 \end{definition}
-It turns out that this is enough for the conflict relation to be irreflexive. (see \cref{fig:full})
+It turns out that this is enough for the conflict relation to be irreflexive. %(see \cref{fig:full})
 \begin{lemma}
@@
 \end{lemma}
-\input{fig/encode_full.tex} 
+%\input{fig/encode_full.tex} 
 So, we see that from the consistent event structure one can get the Prime Event Structure

### 2021-04-01 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: WIP --- reworking text

@@
 $$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge f_{\lPO}(e_1) = f_{\lPO}(e2) $$
+
+Finally, we need a way to reconcile the event structure
+with the states of machine's threads. 
+To do so we use a function $\thrdmap : \eventSet \fun \ThrdState$
+which maps event to a thread state obtained as a result of
+execution of event's side-effect.
+
+Let us consider an example. 
+\todo{}
+
+The rules of operational semantics constructing 
+the event structure are presented on~\cref{fig:eventstruct-sem}.
+
+\input{fig/eventstruct-sem.tex}
 

### 2021-04-02 papers/SYRCoSE/overview.tex
SYRCoSE: WIP --- reworking the text, also add the overview section with the general design principles of the library

\section{Overview}

In this section we sketch the design principles of our library.

We build our mechanization on top of \mathcomp~\cite{Mahboubi-Tassi:MATHCOMP17} library 
using the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq system.
We also use the \emph{small-scale reflection} 
methodology~\cite{Gonthier-Assia:SSR2010, Gonthier-al:SSR2016}, 
a key ingredient of the \ssreflect.

The small-scale reflection approach is based on 
pervasive use of symbolic representation of the proof goal, 
which \emph{relfects} the logical representation. 
The symbolic representation can be manipulated 
by the computational engine of the language, 
allowing the user to automate low-level routine 
proof management by using various decision 
and simplification procedures. 
Whenever the user needs to guide the proof 
it can switch to logical representation
and perform some proof steps manually.

To achieve better automation one is recommended 
to use \emph{decidable} and \emph{computable} procedures
whenever possible.
For example, in the context of our library, 
we encode the binary relations of the event structures
as decidable \texttt{bool}-valued relations, 
\ie $\caOrd, \cfRel : E \fun E \fun \texttt{bool}$,
as opposed to \emph{propositional} 
relations of type $E \fun E \fun \texttt{Prop}$.

We also favor the computational encoding of semantics. 
In the line of the recent related works on mechanization 
of operational semantics~\cite{Xia-al:POPL2019, Letan-al:CPP2020, Affeldt-al:ICMPC2019}, 
we encode the semantics as monadic interpreters.  
This allows us to extract~\cite{Letouzey:CCE2008} 
the semantics as a functional program and run it. 
We believe that the possibility to run the semantics 
is a very useful feature, as it allows 
to debug the formal semantics
and helps to develop better intuition about it.

Finally, we use yet another feature of \mathcomp --- 
\emph{packed classes}~\cite{Garillot-al:ICTPHOL2009}, 
in order to encode the algebraic hierarchy
of various classes of event structures.

### 2021-04-02 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: WIP --- reworking the text, also add the overview section with the general design principles of the library

@@
 
-\begin{definition}[Identifiers set]
+\begin{definition}
   We say that set $\eventSet$ together with strict partial order $\identOrd$
@@
 
+\input{fig/eventstruct-sem.tex}
+
 The rules of operational semantics constructing 
 the event structure are presented on~\cref{fig:eventstruct-sem}.
+The first auxiliary rule \ESAddEventRule adds a new event, sets its 
+label, $\lPO$ and $\lRF$ predecessors. 
+The \ESIdleRule handles the case when a thread of 
+the register machine performs an internal step 
+without any side effect. 
+It chooses an event $e$ together with 
+the thread state $s$ corresponding to it
+and performs one step reduction to new state $s'$.
+It then updates the mapping of events to thread states.   
+The last two rules \ESStoreRule and \ESLoadRule
+correspond to store and load perfromed by some thread.  
+Similarly to \ESIdleRule, an event $e_{\lPO}$ is selected
+and one reduction is performed from the corresponding thread state $s$.
+Unlike the \ESIdleRule case, however, a new event $e$ is also generated.
+In case of \ESLoadRule additionally an event $e_{\lRF}$ is selected,
+such that it has a write label matching the read label of the new event.
 
-\input{fig/eventstruct-sem.tex}
+The following theorem asserts that the event structure built this way
+indeed satisfies the axioms of the prime confusion free event structure.
 
-\eupp{Continue from here}
+\begin{theorem}
+  The tuple $\tup{E, \caOrd, \primcfRel}$, where $\caOrd$ and $\primcfRel$
+  are defined as described above, forms prime confusion-free event structure.
+\end{theorem}
 
-As we saw earlier in Step by Step Example, our event structure contains functions such as $\frf$ and $\fpred$.
-The events themselves are elements of some set $\mathcal{E}$ (set of identifiers), which must satisfy the following properties
-We can fix any $\mathcal{E}$ (for example $\mathbb{N}$) and theat programm's events as a finite subset of $\mathcal{E}$
+%% I can't use `proof` here, sorry.  
+% \begin{proof}
+% \end{proof}
 
-Let $E$ be the set of all program events, then we have
-  $$\ffun{lab} : E \to \lb $$
-where $\lb$ is the label's set (such that $\wlab{x}{5}, \rlab{y}{6}$ ThreadStart, etc.)
-
-We will say that $e$ is a "read" -event with location x and value i if $\ffun{lab}(e) = \rlab{x}{i}$, similarly to "write" -event
-
-As a result, we formally build a four
-\begin{definition}[Execution Event Structure]
-  $$exec\_event\_structure \ \triangleq \ \langle E, \ \ffun{lab}, \ \fpred, \ \frf \rangle$$
-  where
-  \begin{itemize}
-    \item $E$ ~--- finite subset of some identifiers set $\mathcal{E}$
-    \item $\fpred$ ~--- function that returns a predsessor (w.r.t. the program order) of event
-    \item $\frf$ ~--- function that takes "read"-event $e$ and returns a "write"-event from witch $e$ reads
-    \item $\ffun{lab} : E \to \lb$ ~--- function that returns event's label
-  \end{itemize}
-  Also we require
-  \begin{itemize}
-    \item $\forall e \in E, \ \fpred(e) \leqslant_E e$ and $\frf(e) \leqslant_E e$
-    \item $\forall e_1 \ e_2 \in E, \ e_2 = \frf(e_1) \ \Leftrightarrow \ e_1$ is a "read"-event and $e_2$ is a "write"-event with the same location and value
-  \end{itemize}
-\end{definition}
-Further, for simplicity, we will identify identifiers with their labels, and Execution Event Structure with their set of identifiers
-
-Now our goal is to define a prime event structures in these terms. To do this, one need to come up with a causality relation $\leqslant$ and a conflict relation $\#$
-
-Let's start with the causality. In our article, we will assume that the event $e_1$ entails the event $e_2$ if
-\begin{itemize}
-  \item Programm instruction, that corresponds to event $e_1$ states before the programm instruction, that corresponds to $e_2$
-  \item Or $\ e_2$ is the "read"-event and it reads from "write"-event $e_1$
-\end{itemize}
-Or formally:
-\begin{itemize}
-  \item $\fpred(e_2) \ =\ e_1$
-  \item Or $\ \frf(e_2) \ =\ e_1$
-\end{itemize}
-Any function $f$ can be treated as a relation
-  $$ a \ f \ b \ \triangleq\ f(b) \ = \ a$$
-With the notation above, we define
-
-\begin{center}
-\begin{definition}[causality relation]
-  $$\leqslant \ \triangleq \ (\fpred \cup \frf)^* $$
-\end{definition}
-\end{center}
-
-Note that the set $E \subset \mathcal{E}$ already contains the total order $ \leqslant_\mathcal{E} $.
-One can note that the resulting casuality relation is its suborder [TODO:see Well-foundedness and monads]
-
-\begin{center}
-\begin{lemma}
-  If $\leqslant$ is a casuality order of $E \subset \mathcal{E}$ then $\leqslant \ \subset \ \leqslant_\mathcal{E}$
-\end{lemma}
-\end{center}
-
-Now let's define a conflict relation. Looking at the conflict relation in \cref{es_def}, we can establish the following fact:
-\begin{center}
-\begin{lemma}\label{lma_confl}
-  If $\ e_1 \leqslant e_2, \ e_3 \leqslant e_4$ and $e_1 \# e_3$ holds then $e_2 \# e_4$
-\end{lemma} 
-\end{center}
-That is, the conflict relation must be "dragged" through the $\leqslant$ relation.
-Our intuition will be as follows: we, first, define a "minimal" conflict relation, by closing which under the consistency property (\cref{lma_confl}), we finally get a full relation $\#$.
-
-So where do conflicts come from?
-
-From the step-by-step example, we see that if events have a common predsessor (w.r.t. the $\fpred$ relation), then they occur in two different program executions and, accordingly, are conflicting
-\begin{definition}[immediate conflict relation]
-  $$ e_1 \, \ffun{icf} \, e_2 \ \triangleq\ \exists e, \, e = \fpred(e_1) = \fpred(e_2)  $$
-\end{definition}
-Next, one can define the conflict relation as a consistent closure of $\ffun{icf}$
-\begin{definition}[conflict relation]
-  $$ e_1 \, \# \, e_2 \ \triangleq\ \exists e_3\, e_4, e_3 \, \ffun{icf} \, e_4 \ and \ e_3 \leqslant e_1, \, e_4 \leqslant e_2 $$
-\end{definition}
-However, there is one problem with such definition.
-It is clear that the relation defined in this way turns out to be symmetric and consistent.
-Will it be irreflexive?
-It turns out that in the general case, no.
-Consider the following program:
-\input{fig/encode_prog.tex}
-
-Let's build, as in the Step-by-step example, a fragment of $exec\_event\_structure$ corresponding to this program before reading from $y$
-\input{fig/encode_part.tex}
-
-One can observe, that our program in the first thread was divided into two independent executions: execution, in which we read from $x$ value 1 (denote by execution (a)) and execution, in which we read from $x$ 2 (execution (b)).
-
-Now we want to add a read-from-$y$ event to our event structure. Let's see what is going to happen if we will add "read"-event, reading from $\wlab{y}{2}$ (see \cref{fig:cf}).
-\input{fig/encode_cf.tex}
-
-We have $\rlab{x}{1} \leqslant \rlab{y}{2}$, $ \rlab{x}{2} \leqslant \rlab{y}{2}$ and $\rlab{x}{1} \ \# \ \rlab{x}{2}$. Thus $\rlab{y}{2} \ \# \ \rlab{y}{2}$ holds (see~\cref{lma_confl})
-
-To deal with this program, we need to impose a constraint on the event structure.
-
-We will say that the event structure is consistent if there are no reads from conflicting writes. Formally:
-\begin{definition}[consistent Execution Event Structure]
-  $$consitent\langle E, \ffun{lab}, \fpred, \frf \rangle \ \triangleq \ \forall e \in E, \ \neg (e \ \# \ \frf e)$$
-\end{definition}
-It turns out that this is enough for the conflict relation to be irreflexive. %(see \cref{fig:full})
-\begin{lemma}
-  Let $E$ be a consistent Execution Event Structure (denote as $cexec\_event\_structure$). Then it's conflict relation $\#$ is irreflexive
-\end{lemma}
-%\input{fig/encode_full.tex} 
-So, we see that from the consistent event structure one can get the Prime Event Structure
+\todo{Sketch the proof if we'll have space and time}

### 2021-04-05 papers/SYRCoSE/future.tex
SYRCoSE: future work section

\section{Future Work}

There are several directions for future work.

First, we plan to apply our library to a wider range of problems. 
We are going to develop a mechanized semantics of some classical 
languages used to model concurrency, in particular 
the calculus of communicating systems~(CCS)~\cite{Milner:80} and 
$\pi$-calculus~\cite{Milner:99}.      
We also plan to continue our work on 
expressing various relaxed models of shared memory 
in terms of event structures.

Second, we want to cover other classes of event structures in our library, 
in particular bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991}
and stable~\cite{Winskel:82, Winskel:86} event structures.
We plan to use them to develop mechanized \emph{denotational} semantics 
of concurrent languages and relaxed shared memory models.

Finally, we plan to mechanize in Coq classical results 
that connect various classes of event structures  
by isomorphishm of configurations~\cite{Nielsen-al:1981, Boudol-Castellani:1991}. 
It will allow to easily establish the connection between operational and denotational 
semantics of concurrent languages.

### 2021-04-05 papers/SYRCoSE/overview.tex
SYRCoSE: WIP --- related work

@@
 We also favor the computational encoding of semantics. 
-In the line of the recent related works on mechanization 
+Similarly to the recent related works on mechanization 
 of operational semantics~\cite{Xia-al:POPL2019, Letan-al:CPP2020, Affeldt-al:ICMPC2019}, 

### 2021-04-05 papers/SYRCoSE/related.tex
SYRCoSE: WIP --- related work

\section{Related Work}

Event structures were introduced by Winskel to study the semantics of 
calculus of communicating systems~\cite{Winskel:82}. 
Several modification of event structures~\cite{Langerak:91, Boudol-Castellani:1991}
were later proposed to tackle similar problems.  
To the best of our knowledge, there were no attempts 
to mechanize the classical theory of event structures 
in proof assistants.

More recently event structures were applied 
in the context of relaxed memory models~%
\cite{Jeffrey-Riely:LICS16, PichonPharabod-Sewell:POPL16, Chakraborty-Vafeiadis:POPL19, Moiseenko-al:ECOOP20}.
Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
that was accompanied with the mechanization in a proof assistant. 
The authors formalized \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
memory memory model in \coq. However, this memory model uses 
custom variant of event structures, that does not 
obey axioms of any conventional class of 
event structures~\cite{Winskel:82, Langerak:91, Boudol-Castellani:1991}. 
This fact makes it harder to reuse and adapt it 
for other applications of the theory.

### 2021-04-05 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: minor fixes

@@
   Let $\leqslant$ be a partial order. 
-  Then $lessdot$ is covering relation \wrt $\leqslant$ whenever 
+  Then $\lessdot$ is covering relation \wrt $\leqslant$ whenever 
   $x \lessdot y$ is true if and only if $x < y$ and 

### 2021-04-05 papers/SYRCoSE/future.tex

@@
 Finally, we plan to mechanize in Coq classical results 
-that connect various classes of event structures  
-by isomorphishm of configurations~\cite{Nielsen-al:1981, Boudol-Castellani:1991}. 
-It will allow to easily establish the connection between operational and denotational 
+that connect various classes of event structures%  
+~\cite{Nielsen-al:1981, Boudol-Castellani:1991}. 
+It would allow us to easily establish the connection between operational and denotational 
 semantics of concurrent languages.

### 2021-04-05 papers/SYRCoSE/fig/regmachine-eventstruct-prog.tex

\begin{center}
    \begin{tikzpicture}
          \node (wxa) %[left of = mid, node distance = 1.1cm] 
            % at (0,1)
            {$1: \ x:=1$};

          \node (wxb) [right of = wxa, node distance = 1.9cm] 
            % at (1,1)
            {$2: \ r:=x$};

          \node (wxc) [right of = wxb, node distance = 1.9cm] 
            % at (2,1)
            {$3: \ x:=2$};

          \draw ($(wxa.north east)$) -- ($(wxa.south east)$);
          \draw ($(wxb.north west)$) -- ($(wxb.south west)$);
          \draw ($(wxb.north east)$) -- ($(wxb.south east)$);
          \draw ($(wxc.north west)$) -- ($(wxc.south west)$);
    \end{tikzpicture}
    \end{center}

### 2021-04-05 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: add an example of ES construction
SYRCoSE: describe the example

@@
 
-Let us consider an example. 
-\todo{}
+Let us consider an example. Given the program below, 
+our semantics builts the event structure as 
+shown on \cref{fig:regmachine-eventstruct-example}.
+
+\input{fig/regmachine-eventstruct-prog.tex}
+
+\input{fig/regmachine-eventstruct-example.tex}
+
+The construction starts from an initial event structure,
+containing for each thread an event labeled by $\tslab{}$.
+Below the label we depict the corresponding thread state.
+Initially, each event is mapped to an initial thread state, 
+consisting of an instruction pointer pointing to the first 
+instruction to execute, and an initial mapping of registers,
+denoted as $\bot$.
+The first step executes the store instruction from 
+the leftmost thread, and exits the program, 
+since the execution of this thread terminates
+(we omit $\exitInst$ instructions at the end of each thread for brevity).
+Next the store from the rightmost thread is executed and corresponding
+write event is added to the structure. 
+After that the load instruction from the middle thread is executed. 
+Since there are two matching write events in the event structure, 
+two conflicting reads are conjoined to the event structure. 
+Note that the events can be added non-deterministically in any order
+respecting the causality. We could have first execute the rightmost thread
+and add write $\wlab{x}{2}$ before $\wlab{x}{1}$, 
+or add the read with label $\rlab{x}{2}$ 
+before another read $\rlab{x}{1}$.
 

### 2021-04-06 papers/SYRCoSE/future.tex
SYRCoSE: various minor fixes

@@
 We also plan to continue our work on 
-expressing various relaxed models of shared memory 
+expressing various relaxed models 
+of shared memory~\cite{Sewell-al:CACM10, Lahav-al:POPL16, Lahav-al:PLDI17, Podkopaev-al:POPL19}
 in terms of event structures.
@@
 We plan to use them to develop mechanized \emph{denotational} semantics 
-of concurrent languages and relaxed shared memory models.
+of concurrent languages and relaxed shared memory models~\cite{Dodds-al:ESOP18}.
 

### 2021-04-06 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: various minor fixes

@@
 
-In this section we present alternative operational semantics
+In this section we present operational semantics
 which constructs a prime event structure encoding a 
@@
 
-Following the theory of axiomatic weak memory models~\cite{\todo{}},
+We will encode the event structure as a tuple 
+$\tup{\eventSeq, \lLAB, \lfPO, \lfRF}$, 
+below we explain the meaning of each component, 
+and how they together form a prime event structure.
+
+The first component $\eventSeq$ is a sequence of events 
+$e_1 \identOrdDual \dots \identOrdDual e_n$
+in the order reverse to the the one 
+in which events are added to the structure. 
+The second component is labeling function $\lLAB : \eventSet \fun \Lab$, 
+assigning a label to each event.
+
+Next, following the theory of axiomatic weak memory models~\cite{\todo{}},
 we define the causality relation of the register machine's event structure
@@
 In order to construct $\lPO$ and $\lRF$ incrementally
-we represent them via their inverse covering functions.
+we represent them via their inverse covering functions $\lfPO$ and $\lfRF$.
 
@@
 
-  \lessdot_{\lPO} &\defeq& {\frel{f_{\lPO}}}^{-1}  & \lessdot_{\lRF} &\defeq& {\frel{f_{\lRF}}}^{-1} \\
+  \lessdot_{\lPO} &\defeq& {\frel{\lfPO}}^{-1}  & \lessdot_{\lRF} &\defeq& {\frel{\lfRF}}^{-1} \\
   \lPO           &\defeq& \lessdot_{\lPO}^+      & \lRF           &\defeq& \lessdot_{\lRF}       \\
@@
 
-The primitive conflict relation is generated by the $f_{\lPO}$ function.
+The primitive conflict relation is generated by the $\lfPO$ function.
 The two events are considered to be in primitive conflict if they are
@@
 For this definition to work properly, we also need 
-to assume that in the event structure each thread has 
-a special initial event labeled by a distinguished 
-\emph{thread start} label $\tslab{}$.
+to assume that in each thread has a special initial event 
+labeled by a distinguished \emph{thread start} label $\tslab{}$.
 
-$$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge f_{\lPO}(e_1) = f_{\lPO}(e2) $$
+$$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge f_{\lfPO}(e_1) = f_{\lfPO}(e2) $$
 
@@
 with the states of machine's threads. 
-To do so we use a function $\thrdmap : \eventSet \fun \ThrdState$
+To do so we use a function ${\thrdmap : \eventSet \fun \ThrdState}$
 which maps event to a thread state obtained as a result of
@@
 Let us consider an example. Given the program below, 
-our semantics builts the event structure as 
+our semantics builds the event structure as 
 shown on \cref{fig:regmachine-eventstruct-example}.
@@
 the thread state $s$ corresponding to it
-and performs one step reduction to new state $s'$.
+and performs one step reduction to a new state $s'$.
 It then updates the mapping of events to thread states.   

### 2021-04-06 papers/SYRCoSE/regmachine.tex
SYRCoSE: various minor fixes

@@
               & &   & | & \jumpInst{r}{\ip{i}}            & \text{conditional jump}        \\
+              & &   & | & \exitInst                       & \text{exit}                    \\
               & &   & | & \readInst{r}{x}                 & \text{read from memory}        \\
@@
   v      &\in& \Z       &   &                    & \text{value}                   \\
+  \otimes&\in& \BinOp   &   &                    & \text{binary operation}        \\
   \ip{i} &\in& \N       &   &                    & \text{instruction label}       \\
@@
 the semantics of the register machine 
-from a \emph{memory model}~\cite{\todo{}}.
+from a \emph{memory model}.
 
@@
 The semantics is given in a form of 
-\emph{labeled transition system}~\cite{\todo{}}.
+\emph{labeled transition system}~%
+\footnote{As we have mentioned, in our \coq development we actually use 
+the monadic encoding of the operational semantics. 
+The labeled transition system can be deduced from this encoding.}
 State of the system consists of instruction pointer

### 2021-04-06 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: proof sketch

@@
 
-$$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge f_{\lfPO}(e_1) = f_{\lfPO}(e2) $$
+$$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge \lfPO(e_1) = \lfPO(e2) $$
 
@@
 
-%% I can't use `proof` here, sorry.  
-% \begin{proof}
-% \end{proof}
+We sketch the proof below.
 
-\todo{Sketch the proof if we'll have space and time}
+First, we need to show that $\caOrd \ \defeq \ (\lPO \cup \lRF)^*$
+is a partial order. Reflexivity and transitivity follows immediately 
+from the definition of the reflexive-transitive closure. 
+To show antisymmetry note that $\lessdot_{\lPO} \subseteq \identOrd$
+and $\lessdot_{\lRF} \subseteq \identOrd$ by construction. 
+Therefore $\caOrd \subseteq \identOrd$. Since $\identOrd$ is
+partial order, it is antisymmetric, and thus $\caOrd$ should also be antisymmetric.
+The axiom of finite cause, \ie $\prefix{e}$ is finite for every event $e$, 
+follows from the fact that at each step of the construction 
+the set of possible predecessors of the new event 
+can be over-approximated by the finite sequence $\eventSeq$.
+
+Second, we need to show that the conflict relation defined as 
+$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge \lfPO(e_1) = \lfPO(e2) $
+obeys the laws of primitive conflict relation. 
+Trivially, this relation is irreflexive, symmetric, and transitive.
+Also, it is easy to see that it satisfies choice-locality. 
+The side condition $\neg (e_{\lPO} \cfRel e_{\lRF})$ of the rule \ESLoadRule
+ensures that the primitive conflict relation would not violate 
+hereditary property.

### 2021-04-06 papers/SYRCoSE/future.tex
SYRCoSE: minor fixes

@@
 expressing various relaxed models 
-of shared memory~\cite{Sewell-al:CACM10, Lahav-al:POPL16, Lahav-al:PLDI17, Podkopaev-al:POPL19}
+of shared memory~\cite{Lahav-al:POPL16, Lahav-al:PLDI17, Podkopaev-al:POPL19}
 in terms of event structures.

### 2021-04-06 papers/SYRCoSE/intro.tex
SYRCoSE: WIP --- reworking text
SYRCoSE: minor fixes

@@
 to model-based mutation testing~\cite{Fellner-al:VMCAI2020}.
-\eupp{perhaps, we could add more citations once the related work is ready.}
+% \eupp{perhaps, we could add more citations once the related work is ready.}
 
@@
 as depicted on \cref{fig:intro-traces}.
-Event themselves represent atomic side-effects 
-of instructions' execution. In our case 
-it is either a write of a value to a shared variable $\wlab{x}{a}$,
+Event themselves represent atomic side-effects produced by the execution of instruction. 
+In our case it is either a write of a value to a shared variable $\wlab{x}{a}$,
 or a read of a particular value from a shared variable $\rlab{x}{a}$.
@@
 
-In the programming language and formal semantics research community 
+In the programming language and formal semantics research communities 
 it becomes the de-facto standard to complement the theoretical 

### 2021-04-06 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: minor fixes

@@
 
-Next, following the theory of axiomatic weak memory models~\cite{\todo{}},
+Next, following the theory of axiomatic weak memory models~\cite{Lahav-al:PLDI17},
 we define the causality relation of the register machine's event structure
@@
 write events to read events, and ensures that values 
-do not appear out of thin-air~\cite{\todo{}}.
+do not appear out of thin-air~\cite{Boehm-Demsky:MSPC14, Lahav-al:PLDI17}.
 
@@
 hereditary property.
+
+On \cref{fig:eventstruct-example} one can see the prime event structure
+obtained as a result of incremental construction 
+depicted on \cref{fig:regmachine-eventstruct-example}.
+
+\input{fig/eventstruct-example.tex}
+
+Once the event structure is constructed, one can extract
+the configurations corresponding to the particular runs 
+of the parallel register machine, and 
+further filter them via \emph{consistency predicate}
+defining the \emph{memory consistency model}.
+Our construction of the event structures allows to encode 
+a wide class of so-called $\lPO\cup\lRF$ acyclic 
+relaxed memory models~\cite{Lahav-al:PLDI17}.

### 2021-04-08 papers/SYRCoSE/future.tex
Typos and British spelling
find . -name "*.tex" -maxdepth 1 -exec aspell --lang=en_GB --mode=tex check "{}" \;

@@
 First, we plan to apply our library to a wider range of problems. 
-We are going to develop a mechanized semantics of some classical 
+We are going to develop a mechanised semantics of some classical 
 languages used to model concurrency, in particular 
@@
 and stable~\cite{Winskel:82, Winskel:86} event structures.
-We plan to use them to develop mechanized \emph{denotational} semantics 
+We plan to use them to develop mechanised \emph{denotational} semantics 
 of concurrent languages and relaxed shared memory models~\cite{Dodds-al:ESOP18}.
 
-Finally, we plan to mechanize in Coq classical results 
+Finally, we plan to mechanise in Coq classical results 
 that connect various classes of event structures%  

### 2021-04-08 papers/SYRCoSE/intro.tex
Typos and British spelling
find . -name "*.tex" -maxdepth 1 -exec aspell --lang=en_GB --mode=tex check "{}" \;

@@
 compared to traditional interleaving semantics 
-is that they give a more compact and consice 
-representation of progams' behavior.
+is that they give a more compact and concise 
+representation of programs' behaviour.
 For example, consider the following code snippet.
@@
 In the event structure there are two types of edges 
-between the events. The gray arrows $e_1 \arrowCA e_2$ 
+between the events. The grey arrows $e_1 \arrowCA e_2$ 
 represent the \emph{causality relation}, a 
@@
 it becomes the de-facto standard to complement the theoretical 
-studies with their mechanization in the \emph{proof assistants}
+studies with their mechanisation in the \emph{proof assistants}
 like Coq, Agda, Isabelle, Arend, and others,
@@
 Yet, as far as we know, there is a little work on 
-mechanization of event structures theory.
+mechanisation of event structures theory.
 
@@
 and tactics that would allow the researchers 
-to utilize the theory of event structures 
+to utilise the theory of event structures 
 for the needs of their domain.
@@
 In this work in progress report we sketch 
-the common design princples behind our library
+the common design principles behind our library
 and give a concrete example of its usage  
-by developing a formal mechanized semantics of simple 
+by developing a formal mechanised semantics of simple 
 register machine with shared memory.

### 2021-04-08 papers/SYRCoSE/overview.tex
Typos and British spelling
find . -name "*.tex" -maxdepth 1 -exec aspell --lang=en_GB --mode=tex check "{}" \;

@@
 
-We build our mechanization on top of \mathcomp~\cite{Mahboubi-Tassi:MATHCOMP17} library 
+We build our mechanisation on top of \mathcomp~\cite{Mahboubi-Tassi:MATHCOMP17} library 
 using the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq system.
@@
 pervasive use of symbolic representation of the proof goal, 
-which \emph{relfects} the logical representation. 
+which \emph{reflects} the logical representation. 
 The symbolic representation can be manipulated 
@@
 
-We also favor the computational encoding of semantics. 
-Similarly to the recent related works on mechanization 
+We also favour the computational encoding of semantics. 
+Similarly to the recent related works on mechanisation 
 of operational semantics~\cite{Xia-al:POPL2019, Letan-al:CPP2020, Affeldt-al:ICMPC2019}, 

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Typos and British spelling
find . -name "*.tex" -maxdepth 1 -exec aspell --lang=en_GB --mode=tex check "{}" \;

@@
 which constructs a prime event structure encoding a 
-behavior of the register machine.
+behaviour of the register machine.
 
@@
 in which events are added to the structure. 
-The second component is labeling function $\lLAB : \eventSet \fun \Lab$, 
+The second component is labelling function $\lLAB : \eventSet \fun \Lab$, 
 assigning a label to each event.
@@
 We use the inverse covering function because it is 
-more convinient in our setting. Indeed, the semanics adds new event
-at each step. Then it is convinient to require that in addition 
+more convenient in our setting. Indeed, the semantics adds new event
+at each step. Then it is convenient to require that in addition 
 the small-step relation is provided with the $\lPO$ and $\lRF$
@@
 The two events are considered to be in primitive conflict if they are
-not equal and have a commond $\lPO$ predecessor. 
+not equal and have a common $\lPO$ predecessor. 
 For this definition to work properly, we also need 
 to assume that in each thread has a special initial event 
-labeled by a distinguished \emph{thread start} label $\tslab{}$.
+labelled by a distinguished \emph{thread start} label $\tslab{}$.
 
@@
 The construction starts from an initial event structure,
-containing for each thread an event labeled by $\tslab{}$.
+containing for each thread an event labelled by $\tslab{}$.
 Below the label we depict the corresponding thread state.
@@
 The last two rules \ESStoreRule and \ESLoadRule
-correspond to store and load perfromed by some thread.  
+correspond to store and load performed by some thread.  
 Similarly to \ESIdleRule, an event $e_{\lPO}$ is selected

### 2021-04-08 papers/SYRCoSE/regmachine.tex
Typos and British spelling
find . -name "*.tex" -maxdepth 1 -exec aspell --lang=en_GB --mode=tex check "{}" \;

@@
 
-For our case study we use a simple idealized model 
+For our case study we use a simple idealised model 
 of a register machine, which consists of a finite 
@@
 The semantics is given in a form of 
-\emph{labeled transition system}~%
+\emph{labelled transition system}~%
 \footnote{As we have mentioned, in our \coq development we actually use 
 the monadic encoding of the operational semantics. 
-The labeled transition system can be deduced from this encoding.}
+The labelled transition system can be deduced from this encoding.}
 State of the system consists of instruction pointer

### 2021-04-08 papers/SYRCoSE/related.tex
Typos and British spelling
find . -name "*.tex" -maxdepth 1 -exec aspell --lang=en_GB --mode=tex check "{}" \;

@@
 To the best of our knowledge, there were no attempts 
-to mechanize the classical theory of event structures 
+to mechanise the classical theory of event structures 
 in proof assistants.
@@
 Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
-that was accompanied with the mechanization in a proof assistant. 
-The authors formalized \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
+that was accompanied with the mechanisation in a proof assistant. 
+The authors formalised \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
 memory memory model in \coq. However, this memory model uses 

### 2021-04-08 papers/SYRCoSE/intro.tex
Introduction: typos, grammar, wording

@@
 
-The main advantage of the event structures 
+The main advantage of event structures
 compared to traditional interleaving semantics 
 is that they give a more compact and concise 
-representation of programs' behaviour.
-For example, consider the following code snippet.
+representation of programs' behaviours.
+For example, consider the following code snippet
+of a simple parallel program.
 
@@
 it has $3! = 6$ traces with each trace consisting of $4$ events,
-as depicted on \cref{fig:intro-traces}.
-Event themselves represent atomic side-effects produced by the execution of instruction. 
-In our case it is either a write of a value to a shared variable $\wlab{x}{a}$,
-or a read of a particular value from a shared variable $\rlab{x}{a}$.
+as depicted in \cref{fig:intro-traces}.
+Events themselves represent atomic side-effects produced by instruction executions.
+In our case an event is either a \emph{write} of a value $a$ to a shared variable $x$ denoted as $\wlab{x}{a}$,
+or a \emph{read} of a value $a$ from a shared variable $x$ denoted as $\rlab{x}{a}$.
 
@@
 
-In the programming language and formal semantics research communities 
-it becomes the de-facto standard to complement the theoretical 
-studies with their mechanisation in the \emph{proof assistants}
+The programming languages theory and formal semantics research communities 
+are moving to increase the usage of \emph{proof assistants} 
 like Coq, Agda, Isabelle, Arend, and others,
+to complement theoretical studies with their mechanisation,
 as this process increases the reliability and reproducibility 
-of the scientific results.
-Yet, as far as we know, there is a little work on 
-mechanisation of event structures theory.
+of scientific results.
+Yet, to the best of our knowledge, there is little work on 
+mechanisation of the theory of event structures.
+The present report aims to close the gap.
 
-Our work is aimed to close this gap. 
 Our goal is to develop a Coq library containing 
 a comprehensive set of common definitions, lemmas, 
-and tactics that would allow the researchers 
+and tactics that would allow researchers 
 to utilise the theory of event structures 

### 2021-04-08 papers/SYRCoSE/related.tex
TODO
Related work: typos / grammar

@@
 
+\todo{Fix Winskel1986 vs Winskel 1982 (introduction vs related work)}
+
 Event structures were introduced by Winskel to study the semantics of 
-calculus of communicating systems~\cite{Winskel:82}. 
-Several modification of event structures~\cite{Langerak:91, Boudol-Castellani:1991}
+the calculus of communicating systems~\cite{Winskel:82}. 
+Several modifications of event structures~\cite{Langerak:91, Boudol-Castellani:1991}
 were later proposed to tackle similar problems.  

### 2021-04-08 papers/SYRCoSE/related.tex
Related work: remove what we said in Intro
SYRCoSE: apply @dmitromikh suggestions
Related work: more grammar

@@
 were later proposed to tackle similar problems.
-To the best of our knowledge, there were no attempts 
-to mechanise the classical theory of event structures 
-in proof assistants.
 
@@
 Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
-that was accompanied with the mechanisation in a proof assistant. 
+that was accompanied with a mechanisation in a proof assistant. 
 The authors formalised \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
 memory memory model in \coq. However, this memory model uses 
-custom variant of event structures, that does not 
+a custom variant of event structures, that does not 
 obey axioms of any conventional class of 
@@
 This fact makes it harder to reuse and adapt it 
-for other applications of the theory.
+to other applications of the theory.

### 2021-04-08 papers/SYRCoSE/related.tex

@@
 The authors formalised \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
-memory memory model in \coq. However, this memory model uses 
-a custom variant of event structures, that does not 
+memory model in \coq. However, this memory model uses 
+custom variant of event structures, that does not 
 obey axioms of any conventional class of 

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: apply @dmitromikh suggestions
SYRCoSE: fix grammar

@@
 We use the inverse covering function because it is 
-more convenient in our setting. Indeed, the semantics adds new event
+more convenient in our setting. Indeed, the semantics adds a new event
 at each step. Then it is convenient to require that in addition 
@@
 For this definition to work properly, we also need 
-to assume that in each thread has a special initial event 
+to assume that each thread has a special initial event 
 labelled by a distinguished \emph{thread start} label $\tslab{}$.

### 2021-04-08 papers/SYRCoSE/related.tex
SYRCoSE: fix TODOs

@@
 
-\todo{Fix Winskel1986 vs Winskel 1982 (introduction vs related work)}
-
 Event structures were introduced by Winskel to study the semantics of 
-the calculus of communicating systems~\cite{Winskel:82}. 
+the calculus of communicating systems~\cite{Winskel:86, Winskel:82}. 
 Several modifications of event structures~\cite{Langerak:91, Boudol-Castellani:1991}

### 2021-04-08 papers/SYRCoSE/overview.tex

@@
 
-We build our mechanisation on top of \mathcomp~\cite{Mahboubi-Tassi:MATHCOMP17} library 
+We build our mechanisation on top of the \mathcomp~\cite{Mahboubi-Tassi:MATHCOMP17} library 
 using the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq system.
@@
 The small-scale reflection approach is based on 
-pervasive use of symbolic representation of the proof goal, 
-which \emph{reflects} the logical representation. 
+pervasive use of symbolic representations \emph{intermixed}
+with logical ones within the confines of the same proof goal,
+as opposed to large-scale reflection which does not allow such mixing.
+Symbolic representations are connected to the corresponding logical ones
+via user-defined \emph{reflect predicates}.
 The symbolic representation can be manipulated 
@@
 
-To achieve better automation one is recommended 
+To achieve better automation one is encouraged 
 to use \emph{decidable} and \emph{computable} procedures

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
"on fig" -> "in fig"

@@
 our semantics builds the event structure as 
-shown on \cref{fig:regmachine-eventstruct-example}.
+shown in \cref{fig:regmachine-eventstruct-example}.
 
@@
 
-On \cref{fig:eventstruct-example} one can see the prime event structure
+In \cref{fig:eventstruct-example} one can see the prime event structure
 obtained as a result of incremental construction 
-depicted on \cref{fig:regmachine-eventstruct-example}.
+depicted in \cref{fig:regmachine-eventstruct-example}.
 

### 2021-04-08 papers/SYRCoSE/intro.tex
SYRCoSE: minor

@@
 
-In this work in progress report we sketch 
+In this work-in-progress report we sketch 
 the common design principles behind our library

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex

@@
   \begin{itemize}
-    \item there exist distinguished initial identifier $e_0 \in \eventSet$;
-    \item there exist function $\fresh : \eventSet \fun \eventSet$ which 
+    \item there exists a distinguished initial identifier $e_0 \in \eventSet$;
+    \item there exists a function $\fresh : \eventSet \fun \eventSet$ which 
       generates a new fresh identifier, \sth

### 2021-04-08 papers/SYRCoSE/regmachine.tex
Regmachine: small fixes

@@
 The syntax of the machine's language is 
-shown on~\cref{fig:regmachine-syntax}.
+shown in~\cref{fig:regmachine-syntax}.
 
@@
 We first present the semantics of a single thread program.
-In this semantics memory access instructions 
-does not actually operate on shared memory,
+Under this semantics, memory access instructions 
+do not operate on shared memory
 but rather produce a \emph{label}
@@
 
-The semantics is given in a form of 
+The semantics is given in the form of a
 \emph{labelled transition system}~%
@@
 the monadic encoding of the operational semantics. 
-The labelled transition system can be deduced from this encoding.}
+The labelled transition system can be derived from this encoding.}
 State of the system consists of instruction pointer

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Regmachine: wording, typos

@@
 \begin{definition}
-  We say that set $\eventSet$ together with strict partial order $\identOrd$
+  We say that a set $\eventSet$ together with strict partial order $\identOrd$
   form an identifier set if:
@@
 We will encode the event structure as a tuple 
-$\tup{\eventSeq, \lLAB, \lfPO, \lfRF}$, 
-below we explain the meaning of each component, 
-and how they together form a prime event structure.
+$\tup{\eventSeq, \lLAB, \lfPO, \lfRF}$ and explain
+below the meaning of each component, 
+and how they form a prime event structure together.
 

### 2021-04-08 papers/SYRCoSE/regmachine.tex
Regmachine: wording, typos

@@
 the monadic encoding of the operational semantics. 
-The labelled transition system can be derived from this encoding.}
-State of the system consists of instruction pointer
+The labelled transition system can be derived from this encoding.}.
+The state of the system consists of an instruction pointer
 and a map from registers to their values, 
@@
   \ip{i}     &\in& \N            &   &                    & \text{instruction pointer} \\
-  \regmap    &\in& \Reg \fun \Z  &   &                    & \text{registers mapping}   \\
+  \regmap    &\in& \Reg \fun \Z  &   &                    & \text{register mapping}   \\
 

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Regmachine: some more love

@@
   We say that a set $\eventSet$ together with strict partial order $\identOrd$
-  form an identifier set if:
+  form an \emph{identifier set} if:
   \begin{itemize}
@@
 $e_1 \identOrdDual \dots \identOrdDual e_n$
-in the order reverse to the the one 
-in which events are added to the structure. 
-The second component is labelling function $\lLAB : \eventSet \fun \Lab$, 
+in reverse order w.r.t the order 
+in which events get added to the structure. 
+The second component is a labelling function $\lLAB : \eventSet \fun \Lab$, 
 assigning a label to each event.

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Regmachine: 'transitive closure' -> 'reflexive transitive closure'

@@
 we define the causality relation of the register machine's event structure
-as a transitive closure of two relations~---~\emph{program order} 
+as the reflexive transitive closure of two relations~---~\emph{program order} 
 and \emph{reads-from}, denoted as $\lPO$ and $\lRF$ correspondingly.

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Regmachine: union of two relations

--- 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
@@
 we define the causality relation of the register machine's event structure
-as a transitive closure of two relations~---~\emph{program order} 
+as the reflexive transitive closure of the union of two relations~---~\emph{program order} 
 and \emph{reads-from}, denoted as $\lPO$ and $\lRF$ correspondingly.

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Regmachine: a bit more 'the'

@@
 
-Program order tracks precedence of events within a single thread. 
-Reads-from relation captures the flow of values from 
+The program order relation tracks precedence of events within a single thread. 
+The reads-from relation captures the flow of values from 
 write events to read events, and ensures that values 

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Regmachine: hopefully I got the articles right

@@
 write events to read events, and ensures that values 
-do not appear out of thin-air~\cite{Boehm-Demsky:MSPC14, Lahav-al:PLDI17}.
+do not appear out of thin air~\cite{Boehm-Demsky:MSPC14, Lahav-al:PLDI17}.
 
@@
 To do so we use a function ${\thrdmap : \eventSet \fun \ThrdState}$
-which maps event to a thread state obtained as a result of
-execution of event's side-effect.
+which maps an event to a thread state obtained as the result of
+the execution of the event's side-effect.
 

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Regmachine: more

@@
 Below the label we depict the corresponding thread state.
-Initially, each event is mapped to an initial thread state, 
+Initially, each event is mapped to an initial thread state
 consisting of an instruction pointer pointing to the first 
-instruction to execute, and an initial mapping of registers,
+instruction to be executed and an initial mapping of registers
 denoted as $\bot$.
 The first step executes the store instruction from 
-the leftmost thread, and exits the program, 
+the leftmost thread and exits the program, 
 since the execution of this thread terminates
-(we omit $\exitInst$ instructions at the end of each thread for brevity).
-Next the store from the rightmost thread is executed and corresponding
-write event is added to the structure. 
+(we omit the $\exitInst$ instructions at the end of each thread for brevity).
+Next, the store from the rightmost thread is executed and corresponding
+write event gets added to the structure. 
 After that the load instruction from the middle thread is executed. 
@@
 Note that the events can be added non-deterministically in any order
-respecting the causality. We could have first execute the rightmost thread
-and add write $\wlab{x}{2}$ before $\wlab{x}{1}$, 
-or add the read with label $\rlab{x}{2}$ 
+respecting causality. We could have first executed the rightmost thread
+and added write $\wlab{x}{2}$ before $\wlab{x}{1}$, 
+or we could have added the read with label $\rlab{x}{2}$ 
 before another read $\rlab{x}{1}$.

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex

@@
   there is no $z$ \sth $x < z$ and $z < y$.
-  A (non-deterministic) function $f : A \fun \pwset{A}$ is a covering function if
+  A (non-deterministic) function $f$ from $A$ to the set of finite subsets of $A$ is a covering function if
   its corresponding relation, \ie ${\frel{f} \defeq \set{\tup{x, y} ~|~ y \in f(x)}}$, 

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
Regmachine: proof-reading done

@@
 The following theorem asserts that the event structure built this way
-indeed satisfies the axioms of the prime confusion free event structure.
+indeed satisfies the axioms of the prime confusion-free event structure.
 
@@
 Therefore $\caOrd \subseteq \identOrd$. Since $\identOrd$ is
-partial order, it is antisymmetric, and thus $\caOrd$ should also be antisymmetric.
+a partial order, it is antisymmetric, and thus $\caOrd$ should also be antisymmetric.
 The axiom of finite cause, \ie $\prefix{e}$ is finite for every event $e$, 
@@
 ensures that the primitive conflict relation would not violate 
-hereditary property.
+the hereditary property.
 
 In \cref{fig:eventstruct-example} one can see the prime event structure
-obtained as a result of incremental construction 
+obtained as a result of the incremental construction 
 depicted in \cref{fig:regmachine-eventstruct-example}.
@@
 of the parallel register machine, and 
-further filter them via \emph{consistency predicate}
+further filter them via the \emph{consistency predicate}
 defining the \emph{memory consistency model}.
-Our construction of the event structures allows to encode 
+Our construction of event structures allows to encode 
 a wide class of so-called $\lPO\cup\lRF$ acyclic 

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: do not mention confusion-free ES

--- 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex
@@
 
-The primitive conflict relation is generated by the $\lfPO$ function.
+We define the conflict relation in two steps. 
+First, we define the primitive conflict relation $\primcfRel$ 
+which is generated by the $\lfPO$ function.
 The two events are considered to be in primitive conflict if they are
@@
 
-Finally, we need a way to reconcile the event structure
+Second, we extend the primitive conflict along the causality relation:
+
+$$ e_1 \cfRel e_2 \iff 
+   \exists e'_1, e'_2 \in \eventSet \ldotp e'_1 \primcfRel e'_2 \land
+   e'_1 \caOrd e_1 \land e'_2 \caOrd e_2 $$
+
+We also need a way to reconcile the event structure
 with the states of machine's threads. 
@@
 The following theorem asserts that the event structure built this way
-indeed satisfies the axioms of the prime confusion free event structure.
+indeed satisfies the axioms of the prime event structure.
 
@@
   The tuple $\tup{E, \caOrd, \primcfRel}$, where $\caOrd$ and $\primcfRel$
-  are defined as described above, forms prime confusion-free event structure.
+  are defined as described above, forms prime event structure.
 \end{theorem}
@@
 
-Second, we need to show that the conflict relation defined as 
-$ e_1 \primcfRel e_2 \iff e_1 \neq e_2 \wedge \lfPO(e_1) = \lfPO(e2) $
-obeys the laws of primitive conflict relation. 
-Trivially, this relation is irreflexive, symmetric, and transitive.
-Also, it is easy to see that it satisfies choice-locality. 
+Second, we need to show that the conflict relation $\cfRel$ defined as 
+described above obeys the laws of the conflict relation. 
+Trivially, this relation is symmetric, and obeys the hereditary property.
 The side condition $\neg (e_{\lPO} \cfRel e_{\lRF})$ of the rule \ESLoadRule
-ensures that the primitive conflict relation would not violate 
-hereditary property.
+ensures that the conflict relation is irreflexive.
 

### 2021-04-08 papers/SYRCoSE/future.tex

@@
 First, we plan to apply our library to a wider range of problems. 
-We are going to develop a mechanised semantics of some classical 
+We are going to develop a mechanised semantics of some long-established
 languages used to model concurrency, in particular 

### 2021-04-08 papers/SYRCoSE/regmachine-eventstruct.tex

@@
 In \cref{fig:eventstruct-example} one can see the prime event structure
-obtained as a result of incremental construction 
+obtained as a result of the incremental construction 
 depicted in \cref{fig:regmachine-eventstruct-example}.

### 2021-04-15 papers/SYRCoSE/intro.tex
Isabelle -> Isabelle/HOL
Isabelle is a framework, not a proof assistant per se

@@
 are moving to increase the usage of \emph{proof assistants} 
-like Coq, Agda, Isabelle, Arend, and others,
+like Coq, Agda, Isabelle/HOL, Arend, and others,
 to complement theoretical studies with their mechanisation,

### 2021-04-15 papers/SYRCoSE/overview.tex
fix: grammar
clarify perks of using decidable thigs

@@
 methodology~\cite{Gonthier-Assia:SSR2010, Gonthier-al:SSR2016}, 
-a key ingredient of the \ssreflect.
+a key ingredient of \ssreflect.
 
@@
 
-To achieve better automation one is encouraged 
+To achieve better automation and e.g. get \emph{proof irrelevance}
+automatically, one is encouraged
 to use \emph{decidable} and \emph{computable} procedures

### 2021-04-15 papers/SYRCoSE/overview.tex
add TODO for formalization part
add: finmap + relation-algebra

@@
 We build our mechanisation on top of the \mathcomp~\cite{Mahboubi-Tassi:MATHCOMP17} library 
-using the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq system.
+using the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq~system.
 We also use the \emph{small-scale reflection} 
@@
 
-Finally, we use yet another feature of \mathcomp --- 
+To encode the algebraic hierarchy of various classes of event structures
+we use yet another feature of \mathcomp~--- 
 \emph{packed classes}~\cite{Garillot-al:ICTPHOL2009},
-in order to encode the algebraic hierarchy
-of various classes of event structures.
+which is a design pattern providing multiple inheritance,
+maximal sharing of notations and theories,
+automated structure inference and it also deals with performance issues.
+
+To facilitate computable semantics, we define a subclass of
+finite execution event structures as a sorted sequence of events
+combined with finitely supported functions to enhance events with labels,
+event predecessors and a functional version of the reads-from relation.
+
+Encoding finitely supported functions is not a trivial endeavor in
+a proof assistant and for this use the \finmap~%
+\footnote{\url{https://github.com/math-comp/finmap}}
+library which is an extension of \mathcomp~providing finite sets and
+finite maps on types with a choice opeartor (rather than on finite types).
+
+Since our formalization utilizes relations heavily, we have chosen
+to rely on the \relationalgebra~%
+\footnote{\url{https://github.com/damien-pous/relation-algebra}}
+package which provides a number of tactics to solve goals using
+decision procedures for various kinds of relation algebra and
+also lattice theory.
+
+% TODO
+% The relation-algebra package and lattice theory
+% well-foundedness and transitive closure for boolean relations
+% identType and how to fight quadratic time complexity

### 2021-04-15 papers/SYRCoSE/overview.tex
overview: minor

--- 2021-04-15 papers/SYRCoSE/overview.tex
@@
 
+In fact, binary relations are omnipresent in our formalization.
+In order to further automate the reasoning about them, 
+we have chosen to use the \relationalgebra%
+\footnote{\url{https://github.com/damien-pous/relation-algebra}}~%
+\cite{Pous-ITP2013} package 
+which provides a number of tactics to solve goals using
+decision procedures for the theory of lattices  
+and \emph{Kleene Algebras with Tests}~\cite{Kozen:TOPLAS:1997}.
+
 We also favour the computational encoding of semantics. 
@@
 
-Finally, we use yet another feature of \mathcomp --- 
+To facilitate computable semantics, we define a subclass of
+finitely supported event structures as a finite sequence of events
+combined with finitely supported function which enhance events with 
+additional information, such as their labels, causality predecessors, \etc 
+Encoding finitely supported functions is not a trivial endeavor in
+a proof assistant and for this task we use the \finmap~%
+\footnote{\url{https://github.com/math-comp/finmap}}
+library which is an extension of \mathcomp~providing finite sets and
+finite maps on types with a choice opeartor (rather than on finite types).
+
+Finally, to encode the algebraic hierarchy of various classes of event structures
+we use yet another feature of \mathcomp~--- 
 \emph{packed classes}~\cite{Garillot-al:ICTPHOL2009},
-in order to encode the algebraic hierarchy
-of various classes of event structures.
+which is a design pattern providing multiple inheritance,
+maximal sharing of notations and theories,
+and automated structure inference. % and it also deals with performance issues.
+
+% TODO
+% The relation-algebra package and lattice theory
+% well-foundedness and transitive closure for boolean relations
+% identType and how to fight quadratic time complexity

### 2021-05-05 papers/SYRCoSE/response.txt
SYRCoSE: response

We would like to thank the reviewers for their valuable comments. 
We also thank the Reviewer #2 for the reference to the interesting paper relevant to our work.
We proceed with answering the Reviewer #3 questions.

> The Coq proof assistant should be correctly cited, not only particular Coq libraries

We will cite the relevant papers in camera-ready version of the paper.

> Unfortunately, the Event Structure (ES) of Fig2 does not correspond exactly to the set of executions of Fig1: Fig1 suggests that the read operation is carried out after all write actions; this cannot be concluded from the ES of Fig2.

> Fig 9: the prime event structure forgets the difference between the green and the black preorder, therefore it does not express the requirement that in order R(x,2) to occur it must be "immediately preceded" by W(x,2) (in the sense "last write of x before the read"). If you consider this as a prime event structure an execution W(x,2) . W(x,1) . R(x,2) is not forbidden, as W(1,x) and W(2,x) are independent

We answer both of this questions below, as they are connected to each other.

The requirement that the "read is preceded by the last write to the same location" 
in fact corresponds to the sequentially consistent memory model. 
In other models (e.g. total store order TSO, or release/acquire RA) this condition is relaxed
(these models do not require an existance of total order on all events within a single execution (configuration)
and thus in general there is no unique "last" write to read-from).

Our construction of event structure guarantees only the basic condition --- 
acyclicity of program order and reads from. It is, however, a necessary 
precondition for a wide class of memory models, including sequential consistency,
total store order, release/acquire, etc.

Whenever the configuration is extracted from the event structure
it can be checked to additionally satisfy the consistency predicate 
of a particular memory model 
(we briefy mention this in the paper, at the end of Section V).

Moreover, a po+rf acyclic prime event structure together with 
a consistency predicate for a particular memory model 
can be seen as a prime event structure with general (non-binary) conflict. 
In other words, a consistency predicate induces another notion of the conflict.
We are currently working on the theory of prime event structures with 
general conflict that would give us a more solid theoretical background 
for expressing various memory consistency models in the framework of event structures.

As for Fig.1 and Fig.2, you are right here. 
Fig.1. corresponds to sequentially consistent executions of the program,
but the Fig.2 presents po+rf acyclic event structure.
We will try to address this problem and remove the confusion.

> A bit more information about the Coq libraries you use, and why yo use them would be welcome. 
There is space to say a bit more on the approach, may be also for those not "fluent in Coq".
The library is not presented at all, only the application. One would like to get some information about the library itself

We will try to address this problem and give more detailed description
of both our library itself and its dependencies.

> Fig 7: the program example suggests that there is a third behavior of the read thread which runs fully in parallel with the right threads: it should also be possible to "read too early", that is, to observe R(4,\bottom) or get an error ... depending on the semantic choices (the rules of Fig 6 suggest the 1st option, as "errors" are not foreseen)

Yes, you are right. The event structure on Fig.7 is incomplete (i.e. it does not contain the read R(4,\bottom). 
We omit it only for presentation purposes. We will consider to add the missing event to avoid the confusion.

> Fig 8, Store rule: I suppose here e_rf is an error

Yes, it is a typo.

> Theorem 1: did you use Coq to make this proof ? It should be done

Yes, we have a proof in Coq.

### 2021-05-07 papers/SYRCoSE/intro.tex
intro: link to the git repo
minor grammar
intro: minor
*: apply Reviewer #3 suggestions

--- 2021-04-08 papers/SYRCoSE/intro.tex
@@
 
+\input{fig/intro-traces.tex}
+
 Under the interleaving semantics 
@@
 or a \emph{read} of a value $a$ from a shared variable $x$ denoted as $\rlab{x}{a}$.
-
-\input{fig/intro-traces.tex}
 
@@
 as a linearisation of some \emph{configuration}, 
-that is a causally-closed and conflict-free 
-subset of events.
+that is a causally-closed and conflict-free subset of events,  
+which additionally should satisfy constraint 
+that each read is preceded by a matching write.
 
@@
 are moving to increase the usage of \emph{proof assistants} 
-like Coq, Agda, Isabelle, Arend, and others,
+like Coq~\cite{Coq}, Agda~\cite{Agda}, 
+Isabelle/HOL~\cite{Nipkow-al:Isabelle2002}, 
+Arend~\cite{Arend}, and others,
 to complement theoretical studies with their mechanisation,
@@
 register machine with shared memory.
+
+Our library together with the examples of 
+its usage is available online at \url{https://github.com/volodeyka/event-struct}.

### 2021-05-07 papers/SYRCoSE/regmachine-eventstruct.tex
SYRCoSE: minor
*: apply Reviewer #3 suggestions

@@
 which constructs a prime event structure encoding a 
-behaviour of the register machine.
+set of possible behaviours of the register machine.
 
@@
 the event structure are presented on~\cref{fig:eventstruct-sem}.
-The first auxiliary rule \ESAddEventRule adds a new event, sets its 
+The first auxiliary rule \ESAddEventRule~ adds a new event, sets its 
 label, $\lPO$ and $\lRF$ predecessors. 
-The \ESIdleRule handles the case when a thread of 
+The \ESIdleRule~ handles the case when a thread of 
 the register machine performs an internal step 
 without any side effect. 
-It chooses an event $e$ together with 
-the thread state $s$ corresponding to it
-and performs one step reduction to a new state $s'$.
+It chooses an event~$e$ together with 
+the thread state~$s$ corresponding to it
+and performs one step reduction to a new state~$s'$.
 It then updates the mapping of events to thread states.   
-The last two rules \ESStoreRule and \ESLoadRule
+The last three rules \ESStoreRule, \ESLoadRule, and \ESLoadBotRule~
 correspond to store and load performed by some thread.  
 Similarly to \ESIdleRule, an event $e_{\lPO}$ is selected
-and one reduction is performed from the corresponding thread state $s$.
-Unlike the \ESIdleRule case, however, a new event $e$ is also generated.
-In case of \ESLoadRule additionally an event $e_{\lRF}$ is selected,
+and one reduction is performed from the corresponding thread state~$s$.
+Unlike the \ESIdleRule~ case, however, a new event~$e$ is also generated.
+In case of \ESLoadRule~ additionally an event~$e_{\lRF}$ is selected,
 such that it has a write label matching the read label of the new event.    
+The rule \ESLoadBotRule~ corresponds to a case when load 
+is performed ``too early'', before any write to the given location is available.
 
@@
 \begin{theorem}
-  The tuple $\tup{E, \caOrd, \primcfRel}$, where $\caOrd$ and $\primcfRel$
+  The tuple $\tup{E, \caOrd, \cfRel}$, where $\caOrd$ and $\cfRel$
   are defined as described above, forms prime event structure.
@@
 
-We sketch the proof below.
+We sketch the proof below\footnote{
+One can also find mechanized proof in our Coq development.
+}.
 
@@
 relaxed memory models~\cite{Lahav-al:PLDI17}.
+
+For example, a predicate corresponding
+to \emph{sequential consistency}~\cite{Lamport:TC79}
+requires that the causality order
+can be extended to a total order on all events 
+of the configuration, such that for each read 
+event the last preceding write event 
+to the same location has the same value as the read.

### 2021-05-07 papers/SYRCoSE/regmachine.tex
add: finmap + relation-algebra
*: apply Reviewer #3 suggestions

@@
 The semantics is given in the form of a
-\emph{labelled transition system}~%
-\footnote{As we have mentioned, in our \coq development we actually use 
+\emph{labelled transition system}:~%
+\footnote{As we have mentioned, in our \coq~development we actually use 
 the monadic encoding of the operational semantics. 
-The labelled transition system can be derived from this encoding.}.
-The state of the system consists of an instruction pointer
-and a map from registers to their values, 
+The labelled transition system can be derived from this encoding.}
+$P \vdash s \thrdstep{l} s'$,
+where $P$ is a program, $l$ is a label, $s$ and $s'$ are states of the machine. 
+The state of the machine itself consist of an instruction pointer $\ip{i}$
+and a map from registers to their values $\regmap$, 
 as shown in \cref{fig:regmachine-thrdstate}. 

### 2021-05-09 papers/SYRCoSE/future.tex
Use xspace to handle spaces after \newcommands such as \coq

@@
 
-Finally, we plan to mechanise in Coq classical results 
+Finally, we plan to mechanise in \coq classical results 
 that connect various classes of event structures%  

### 2021-05-09 papers/SYRCoSE/intro.tex
Use xspace to handle spaces after \newcommands such as \coq

--- 2021-04-08 papers/SYRCoSE/intro.tex
@@
 as a linearisation of some \emph{configuration}, 
-that is a causally-closed and conflict-free 
-subset of events.
+that is a causally-closed and conflict-free subset of events,  
+which additionally should satisfy the constraint 
+that each read is preceded by a matching write.
 
@@
 are moving to increase the usage of \emph{proof assistants} 
-like Coq, Agda, Isabelle, Arend, and others,
+like \coq~\cite{Coq}, \agda~\cite{Agda}, 
+\isabelle~\cite{Nipkow-al:Isabelle2002}, 
+\arend~\cite{Arend}, and others,
 to complement theoretical studies with their mechanisation,
@@
 
-Our goal is to develop a Coq library containing 
+Our goal is to develop a \coq library containing 
 a comprehensive set of common definitions, lemmas, 

### 2021-05-09 papers/SYRCoSE/overview.tex
Use xspace to handle spaces after \newcommands such as \coq

@@
 We build our mechanisation on top of the \mathcomp~\cite{Mahboubi-Tassi:MATHCOMP17} library 
-using the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq~system.
+using the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq system.
 We also use the \emph{small-scale reflection} 
@@
 \footnote{\url{https://github.com/math-comp/finmap}}
-library which is an extension of \mathcomp~providing finite sets and
+library which is an extension of \mathcomp providing finite sets and
 finite maps on types with a choice opeartor (rather than on finite types).
@@
 Finally, to encode the algebraic hierarchy of various classes of event structures
-we use yet another feature of \mathcomp~--- 
+we use yet another feature of \mathcomp --- 
 \emph{packed classes}~\cite{Garillot-al:ICTPHOL2009},

### 2021-05-09 papers/SYRCoSE/intro.tex
Typos and grammar mostly

@@
 
-Under the interleaving semantics 
+Under the interleaving semantics,
 it has $3! = 6$ traces with each trace consisting of $4$ events,
@@
 Events themselves represent atomic side-effects produced by instruction executions.
-In our case an event is either a \emph{write} of a value $a$ to a shared variable $x$ denoted as $\wlab{x}{a}$,
+In our case, an event is either a \emph{write} of a value $a$ to a shared variable $x$ denoted as $\wlab{x}{a}$,
 or a \emph{read} of a value $a$ from a shared variable $x$ denoted as $\rlab{x}{a}$.
@@
 (see \cref{fig:intro-es}). 
-In the event structure there are two types of edges 
+In the event structure, there are two types of edges 
 between the events. The grey arrows $e_1 \arrowCA e_2$ 
@@
 
-In this work-in-progress report we sketch 
+In this work-in-progress report, we sketch 
 the common design principles behind our library
 and give a concrete example of its usage  
-by developing a formal mechanised semantics of simple 
+by developing a formal mechanised semantics of a simple 
 register machine with shared memory.

### 2021-05-09 papers/SYRCoSE/overview.tex
Typos and grammar mostly

@@
 
-In this section we sketch the design principles of our library.
+In this section, we sketch the design principles of our library.
 
@@
 The small-scale reflection approach is based on 
-pervasive use of symbolic representations \emph{intermixed}
+the pervasive use of symbolic representations \emph{intermixed}
 with logical ones within the confines of the same proof goal,
@@
 Whenever the user needs to guide the proof 
-it can switch to logical representation
+they can switch to the logical representation
 and perform some proof steps manually.
@@
 automatically, one is encouraged
-to use \emph{decidable} and \emph{computable} procedures
-whenever possible.
+to use \emph{decision} procedures whenever possible.
 For example, in the context of our library, 
@@
 We also favour the computational encoding of semantics. 
-Similarly to the recent related works on mechanisation 
+Similar to the recent related works on mechanisation 
 of operational semantics~\cite{Xia-al:POPL2019, Letan-al:CPP2020, Affeldt-al:ICMPC2019}, 
@@
 finitely supported event structures as a finite sequence of events
-combined with finitely supported function which enhance events with 
+combined with finitely supported function which enhances events with 
 additional information, such as their labels, causality predecessors, \etc 
@@
 library which is an extension of \mathcomp providing finite sets and
-finite maps on types with a choice opeartor (rather than on finite types).
+finite maps on types with a choice operator (rather than on finite types).
 

### 2021-05-09 papers/SYRCoSE/regmachine.tex
Use xspace to handle spaces after \newcommands such as \coq
Typos and grammar mostly

@@
 
-For our case study we use a simple idealised model 
+For our case study, we use a simple idealised model 
 of a register machine, which consists of a finite 
-sequence of instructions, an instruction pointer 
+sequence of instructions, an instruction pointer,
 and an infinite set of registers. 
@@
 \emph{labelled transition system}:~%
-\footnote{As we have mentioned, in our \coq~development we actually use 
+\footnote{As we have mentioned, in our \coq development we actually use 
 the monadic encoding of the operational semantics. 
@@
 where $P$ is a program, $l$ is a label, $s$ and $s'$ are states of the machine. 
-The state of the machine itself consist of an instruction pointer $\ip{i}$
+The state of the machine itself consists of an instruction pointer $\ip{i}$
 and a map from registers to their values $\regmap$, 

### 2021-05-09 papers/SYRCoSE/related.tex
Typos and grammar mostly

@@
 Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
-that was accompanied with a mechanisation in a proof assistant. 
+that was accompanied by a mechanisation in a proof assistant. 
 The authors formalised \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
 memory model in \coq. However, this memory model uses 
-custom variant of event structures, that does not 
+a custom variant of event structures, that does not 
 obey axioms of any conventional class of 

### 2021-05-09 papers/SYRCoSE/intro.tex
more grammar

@@
 by Winskel~\cite{Winskel:86} as a semantic domain of concurrent programs.
-In recent years there is a renewed interest in event structures, 
+In recent years there has been renewed interest in event structures, 
 with the applications of the theory ranging from relaxed memory models%

### 2021-05-09 papers/SYRCoSE/related.tex
Comma for introductory words (but I am not entirely sure here)
related.tex: the

@@
 
-More recently event structures were applied 
+More recently, event structures were applied 
 in the context of relaxed memory models~%
@@
 that was accompanied by a mechanisation in a proof assistant. 
-The authors formalised \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
+The authors formalised the \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
 memory model in \coq. However, this memory model uses 

### 2021-05-09 papers/SYRCoSE/regmachine-eventstruct.tex
Use xspace to handle spaces after \newcommands such as \coq
Move intro words to the end to avoid more commas

@@
 containing for each thread an event labelled by $\tslab{}$.
-Below the label we depict the corresponding thread state.
+We depict the corresponding thread state below the label.
 Initially, each event is mapped to an initial thread state
@@
 We sketch the proof below\footnote{
-One can also find mechanized proof in our Coq development.
+One can also find mechanized proof in our \coq development.
 }.

### 2021-05-09 papers/SYRCoSE/overview.tex
overview.tex: article

@@
 finitely supported event structures as a finite sequence of events
-combined with finitely supported function which enhances events with 
+combined with a finitely supported function which enhances events with 
 additional information, such as their labels, causality predecessors, \etc 

### 2021-05-09 papers/SYRCoSE/regmachine-eventstruct.tex
regmachine-eventstruct.tex: commas and articles

@@
 on each step. In order to generate a new event 
-on each step we require that events behave as \emph{identifiers}.
+on each step, we require that events behave as \emph{identifiers}.
 
@@
 more convenient in our setting. Indeed, the semantics adds a new event
-at each step. Then it is convenient to require that in addition 
+at each step. Then it is convenient to require that, in addition,
 the small-step relation is provided with the $\lPO$ and $\lRF$
@@
 We also need a way to reconcile the event structure
-with the states of machine's threads. 
-To do so we use a function ${\thrdmap : \eventSet \fun \ThrdState}$
+with the states of the machine's threads. 
+To do so, we use a function ${\thrdmap : \eventSet \fun \ThrdState}$
 which maps an event to a thread state obtained as the result of

### 2021-05-09 papers/SYRCoSE/regmachine-eventstruct.tex
minor clarification
commas? (this looks better but still feels a bit weird)

@@
 Let us consider an example. Given the program below, 
-our semantics builds the event structure as 
+our semantics builds the corresponding event structure as 
 shown in \cref{fig:regmachine-eventstruct-example}.
@@
 
-The construction starts from an initial event structure,
-containing for each thread an event labelled by $\tslab{}$.
+The construction starts from an initial event structure
+containing, for each thread, an event labelled by $\tslab{}$.
 We depict the corresponding thread state below the label.

### 2021-05-09 papers/SYRCoSE/regmachine-eventstruct.tex
grammar

@@
 containing, for each thread, an event labelled by $\tslab{}$.
-We depict the corresponding thread state below the label.
+We depict the corresponding thread state below each label.
 Initially, each event is mapped to an initial thread state
@@
 (we omit the $\exitInst$ instructions at the end of each thread for brevity).
-Next, the store from the rightmost thread is executed and corresponding
+Next, the store from the rightmost thread is executed and the corresponding
 write event gets added to the structure. 
-After that the load instruction from the middle thread is executed. 
+After that, the load instruction from the middle thread is executed. 
 Since there are two matching write events in the event structure, 
@@
 The rules of operational semantics constructing 
-the event structure are presented on~\cref{fig:eventstruct-sem}.
+the event structure are presented in~\cref{fig:eventstruct-sem}.
 The first auxiliary rule \ESAddEventRule~ adds a new event, sets its 

### 2021-05-09 papers/SYRCoSE/regmachine-eventstruct.tex
xspace for semantics rules

@@
 the event structure are presented in~\cref{fig:eventstruct-sem}.
-The first auxiliary rule \ESAddEventRule~ adds a new event, sets its 
+The first auxiliary rule \ESAddEventRule adds a new event, sets its 
 label, $\lPO$ and $\lRF$ predecessors. 
-The \ESIdleRule~ handles the case when a thread of 
+The \ESIdleRule handles the case when a thread of 
 the register machine performs an internal step 
@@
 It then updates the mapping of events to thread states.   
-The last three rules \ESStoreRule, \ESLoadRule, and \ESLoadBotRule~
+The last three rules \ESStoreRule, \ESLoadRule, and \ESLoadBotRule
 correspond to store and load performed by some thread.  
@@
 and one reduction is performed from the corresponding thread state~$s$.
-Unlike the \ESIdleRule~ case, however, a new event~$e$ is also generated.
-In case of \ESLoadRule~ additionally an event~$e_{\lRF}$ is selected,
+Unlike the \ESIdleRule case, however, a new event~$e$ is also generated.
+In case of \ESLoadRule additionally an event~$e_{\lRF}$ is selected,
 such that it has a write label matching the read label of the new event.    
-The rule \ESLoadBotRule~ corresponds to a case when load 
+The rule \ESLoadBotRule corresponds to a case when load 
 is performed ``too early'', before any write to the given location is available.

### 2021-05-09 papers/SYRCoSE/regmachine-eventstruct.tex
more grammar

@@
 Unlike the \ESIdleRule case, however, a new event~$e$ is also generated.
-In case of \ESLoadRule additionally an event~$e_{\lRF}$ is selected,
+In the case of \ESLoadRule, additionally, an event~$e_{\lRF}$ is selected,
 such that it has a write label matching the read label of the new event.    
@@
 We sketch the proof below\footnote{
-One can also find mechanized proof in our \coq development.
+One can also find its mechanized proof in our \coq development.
 }.

### 2021-05-09 papers/SYRCoSE/response.txt
fix typos in papers/SYRCoSE/response.txt

@@
 In other models (e.g. total store order TSO, or release/acquire RA) this condition is relaxed
-(these models do not require an existance of total order on all events within a single execution (configuration)
+(these models do not require an existence of total order on all events within a single execution (configuration)
 and thus in general there is no unique "last" write to read-from).
@@
 of a particular memory model 
-(we briefy mention this in the paper, at the end of Section V).
+(we briefly mention this in the paper, at the end of Section V).
 

### 2021-05-09 papers/SYRCoSE/response.txt
fix formatting in papers/SYRCoSE/response.txt

@@
 > A bit more information about the Coq libraries you use, and why yo use them would be welcome. 
-There is space to say a bit more on the approach, may be also for those not "fluent in Coq".
-The library is not presented at all, only the application. One would like to get some information about the library itself
+> There is space to say a bit more on the approach, may be also for those not "fluent in Coq".
+> The library is not presented at all, only the application. One would like to get some information about the library itself
 

### 2021-05-10 papers/SYRCoSE/overview.tex
overview: add more on mathcomp

@@
-\section{Overview}
+\section{Overview of Our Library}
 
@@
 We build our mechanisation on top of the \mathcomp~\cite{Mahboubi-Tassi:MATHCOMP17} library 
-using the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq system.
+which is an extensive and coherent repository of formalized mathematical theories,
+whose implementation is based on the \ssreflect~\cite{Gonthier-al:SSR2016} extension of the \coq system.
+By using \mathcomp, we draw on the large corpus of already formalised algorithms and mathematical results:
+its core modules feature support for a range of useful data structures,
+e.g. numbers, sequences, finite graphs,
+and also interfaces: types with decidable equality, subtypes, finite types, and so on.
+
 We also use the \emph{small-scale reflection} 
@@
 To achieve better automation and e.g. get \emph{proof irrelevance}
-automatically, one is encouraged
+for free, one is encouraged
 to use \emph{decision} procedures whenever possible.

### 2021-05-10 papers/SYRCoSE/regmachine-eventstruct.tex
fix: side condition in Load rule
fix: proof of Threorem 1

@@
 and $\lessdot_{\lRF} \subseteq \identOrd$ by construction. 
-Therefore $\caOrd \subseteq \identOrd$. Since $\identOrd$ is
+Therefore $\caOrd$ is a subset of the reflexive closure of $\identOrd$. Since $\identOrd$ is
 a partial order, it is antisymmetric, and thus $\caOrd$ should also be antisymmetric.
@@
 Trivially, this relation is symmetric, and obeys the hereditary property.
-The side condition $\neg (e_{\lPO} \cfRel e_{\lRF})$ of the rule \ESLoadRule
+The side condition $\neg (e \cfRel e_{\lRF})$ of the rule \ESLoadRule
 ensures that the conflict relation is irreflexive.

### 2021-05-10 papers/SYRCoSE/overview.tex
expand a bit on relation algebra

@@
 In fact, binary relations are omnipresent in our formalization.
-In order to further automate the reasoning about them, 
+This quickly manifested in a substantial amount of proof overhead and
+we seeked for tools to automate our proofs.
+Since binary relations form a \emph{Kleene Algebra with Tests}~(KAT)~\cite{Kozen:TOPLAS:1997},
 we have chosen to use the \relationalgebra%
@@
 which provides a number of tactics to solve goals using
-decision procedures for the theory of lattices  
-and \emph{Kleene Algebras with Tests}~\cite{Kozen:TOPLAS:1997}.
+decision procedures for a number of theories, such as
+partially ordered monoids, lattices, residuated Kleene allegories and KATs.
 

### 2021-05-10 papers/SYRCoSE/overview.tex
mention Equations plugin

@@
 relations of type $E \fun E \fun \texttt{Prop}$.
+
+Encoding computable relations in \coq,
+especially their (computable) transitive closures,
+can be quite challenging since \coq is a total language and
+its termination checker only understands termination patterns
+going slightly beyond simple structural recursion.
+To make it easier, we employ the \equations function definition plugin~%
+\cite{Sozeau-Mangin:ICFP2019}
+which provides both notations for writing programs by dependent pattern-matching
+and good support for well-founded recursion.
 

### 2021-05-10 papers/SYRCoSE/intro.tex
minor

@@
 
-Our goal is to develop a \coq library containing 
+Our end goal is to develop a \coq library containing 
 a comprehensive set of common definitions, lemmas, 

### 2021-05-10 papers/SYRCoSE/intro.tex
change repo address to current
why coq

@@
 are moving to increase the usage of \emph{proof assistants} 
-like \coq~\cite{Coq}, \agda~\cite{Agda}, 
+like \coq~\cite{Coq},
+%\agda~\cite{Agda}, 
 \isabelle~\cite{Nipkow-al:Isabelle2002}, 
@@
 The present report aims to close the gap.
+
+We have chosen \coq as the proof assistant because it's a mature formal proof management tool
+with a rich ecosystem of libraries, plugins, documentation, and
+existing applications including the certification of properties of programming languages:
+the verified C compiler CompCert~\cite{Leroy-Compcert-CACM:2006},
+the Verified Software Toolchain~\cite{Appel-ESOP2011} for verification of C programs,
+and the Iris framework~\cite{Jung-al:JFP2018} for concurrent separation logic,
+to name a few.
 
@@
 Our library together with the examples of 
-its usage is available online at \url{https://github.com/volodeyka/event-struct}.
+its usage is available online at \url{https://github.com/event-structures/event-struct}.

### 2021-05-21 papers/SYRCoSE/fig/encode_prog.tex
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

* feat: add part of small-step

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}

      \node (iwy) %[left of = mid, node distance = 1.1cm] 
      at (-1.3, 2.5)
      {$x:=1;$};

      \node (irx) [above = -0.5cm of iwy.west, anchor = west] 
      {$r:=x;$};

      \node (iwx) %[right of = mid, node distance = 0.8cm] 
       at (0.5, 2.5)
      {$x:=2;$};

      \node (iwz) %[right of = mid, node distance = 0.8cm] 
      at (-1.3, 1.55)
     {$y:=r;$};

     \node (iww) %[right of = mid, node distance = 0.8cm] 
      at (-1.3, 1.2)
     {$r':=y;$};

      \draw ($ (iwy.south east) + (0.1,-1.3) $) -- ($ (iwy.south east) + (0.1,0.6) $);
      \draw ($ (iwx.south west) + (-0.15,-1.3) $) -- ($ (iwx.south west) - (0.15,-0.6) $);
    \end{tikzpicture}
  \end{center}
\end{figure}

### 2021-05-21 papers/SYRCoSE/future.tex
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

* feat: add part of small-step

@@
 We are going to develop a mechanised semantics of some long-established
-languages used to model concurrency, in particular 
+languages used to model concurrency, in particular,
 the calculus of communicating systems~(CCS)~\cite{Milner:80} and 
@@
 Second, we want to cover other classes of event structures in our library, 
-in particular bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991}
+in particular, bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991},
 and stable~\cite{Winskel:82, Winskel:86} event structures.

### 2021-05-21 papers/SYRCoSE/regmachine.tex
typo: single-threaded
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

* feat:

@@
   r      &\in& \Reg     &   &                    & \text{thread-local register} \\ 
-  x      &\in& \Loc     &   &                    & \text{shared memory location}  \\ 
+  x      &\in& \Loc     &   &                    & \text{shared memory }        \\
+         &   &          &   &                    & \qquad  \text{location}      \\
   v      &\in& \Z       &   &                    & \text{value}                 \\
@@
 \]
+
 \caption{Syntax of the register machine}
@@
 
-We first present the semantics of a single thread program.
+We first present the semantics of a single-threaded program.
 Under this semantics, memory access instructions 

### 2021-05-21 papers/SYRCoSE/related.tex
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

* feat: add part of small-step

@@
 a custom variant of event structures, that does not 
-obey axioms of any conventional class of 
+obey the axioms of any conventional class of 
 event structures~\cite{Winskel:82, Langerak:91, Boudol-Castellani:1991}. 

### 2021-05-21 papers/SYRCoSE/small_step.tex
SYRCoSE: WIP --- reworking text
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

@@
   \end{itemize}
-\input{fig/small_step.tex}
+
 \end{definition}
-Then we define programm configuration:
+Then we define program configuration:
 \begin{definition}[configuration]
@@
 First component of this pair is a state of some thread after execution of $e$'s memory-side-effect.
-Our programm contains several threads, so we need to specify number of thread (where everything is going on) in the second component
+Our program contains several threads, so we need to specify number of thread (where everything is going on) in the second component
   \end{itemize}
 \end{definition}
-So programm configuration describes current Execution Event Structure, where we map each event to corresponding thread state.
+So program configuration describes current Execution Event Structure, where we map each event to corresponding thread state.
 Thus, to make our semantics step we should select some event and make progress in it's thread state.
@@
 \\
-If we have some $p : \texttt{P}$, and $t \in \mathbb{N}$ ~--- thread number, $p[t] :=$ $t$'s thread of programm $p$.
+If we have some $p : \texttt{P}$, and $t \in \mathbb{N}$ ~--- thread number, $p[t] :=$ $t$'s thread of program $p$.
 
@@
 Unfortunately, in the last step we could read from conflicting event $w$.
-So after making $\to$ step, it is necessary to chek if our final $\es'$ is consistent (see small-step rule).
+So after making $\to$ step, it is necessary to check if our final $\es'$ is consistent (see small-step rule).
 After all, we obtain one-step relation $\Rightarrow$.
